---
title: "_master"
author: "Xiaoyi Zhao"
date: "2026-07-17"
output: html_document
---

## Purpose

This is the master file, and only file the users technically need to run 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##### Step 1: 
In order to replicate all the figures and output, you will open this file under the R project so that you have it under your own directory 

##### Step 2: Install all the Packages and load all the libraries
```{r}
# you may comment this out once you installed all the required packages
install.packages(packages <- c(
  "sf",
  "dplyr",
  "ggplot2",
  "tidyr",
  "purrr",
  "ggplot2",
  "scales",
  "tidycensus",
  "showtext",
  "sysfonts",
  "rsvg",
  "ggimage",
  "mapview",
  "ggradar",
  "rnassqs"
)
)


```

```{r}
library(tigris)    # county polygons (only used below to locate each market's county)
library(sf)        # the workhorse: stores and manipulates spatial data
library(dplyr)     # regular data wrangling once things look like data frames
library(ggplot2)
library(mapview)
library(readxl)
library(units)
library(rsvg)
library(ggimage)
library(tidyr)
library(purrr)
library(scales)
library(terra)
library(tidycensus)
library(readr)
library(ggnewscale)
library(ggtext) 
library(showtext)
library(sysfonts)
library(patchwork)
library(tidyverse)
library(ggplot2)
library(mapview)
library(readxl)
library(units)
library(ggradar)
library(grid)
library(tigris)
library(rnassqs)
```


##### Step 3: Clean the dataset 

###### 3a: calcualted the distance from county centroid to the closest foodhub

```{r}
rmarkdown::render("../programs/cl_01_distance.Rmd")

```

###### 3b: calcualted the distance from the county centroid to the closest farmers' market

```{r}
rmarkdown::render("../programs/cl_01_farmersmarket_distance.Rmd")
```

###### 3c: separate the variables needed from the index from FAME

```{r}
rmarkdown::render("../programs/cl_01_fame.Rmd")
```
###### 3d: Compute each county's fruit & nut share = (county fruit & nut sales) / (county total farm-related income receipts), using USDA-NASS data.

**This will require the user to input their own USDA API key, by modifying line 24 in the `cl_02_fruitnut_cnty.Rmd`**
```{r}
rmarkdown::render("../programs/cl_02_fruitnut_cnty.Rmd")
```

###### 3e: merge the datasets togehter
```{r}
rmarkdown::render("../programs/cl_03_merge.Rmd")
```

###### 3f: Build the Index
```{r}
rmarkdown::render("../programs/an_01.Rmd")
```
###### 3f: Build the Index for whatif cases used in the fig_02
```{r}
rmarkdown::render("../programs/an_02_spma_simulation.Rmd")
```

##### Step 4: Build the Visualization

```{r}
rmarkdown::render("../programs/fig_01.Rmd")
```

```{r}
rmarkdown::render("../programs/fig_02.Rmd")
```

**Note: If rendering does not work, you can also open the file and run the file under the R proj**
