---
title: "an_01b_spma_index"
author: "Pragati Dahal"
date: "2026-07-15"
output: html_document
---
## Purpose

Compute the Small-Producer Market Access (SPMA) index for every county from
`distance_SPMA_merged.csv`, then map it as a US county choropleth.

## The SPMA index

For county \(i\), let

- \(d^{\text{FM}}_i\) = distance from the county centroid to the nearest **farmers market** (miles),
- \(d^{\text{hub}}_i\) = distance from the county centroid to the nearest **food hub** (miles),

and define the three sales pieces from the county's total sales `all_sales`:

$$
\text{DTC}_i = \text{all_sales}_i \times \text{d2c_sales_pct}_i/100,\qquad
\text{Int}_i = \text{all_sales}_i \times \text{intermediated_sales_pct}_i/100 .
$$

Then

$$
\text{SPMA}_i =
\dfrac{\text{DTC}_i}{\left(d^{\text{FM}}_i\right)^2} \;+\; \dfrac{\text{Int}_i}{\left(d^{\text{hub}}_i\right)^2}
$$

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Load the library
```{r}
library(readr)
library(dplyr)
library(sf)
library(tigris)
library(ggplot2)
library(readxl)
library(ggnewscale)
options(tigris_use_cache = TRUE)
```

### Step 1 — read the merged county table
```{r}
dat <- read_csv("../data/outcome/distance_SPMA_merged.csv") %>%
  mutate(county_fips = sprintf("%05d", as.integer(county_fips)))
```


### Step 2 — derive intermediated sales share from DTC share
`d2c_sales_pct` and `all_sales` have the fewest missing values of any of the sales
variables, so rather than relying on the sparser `intermediated_sales_pct` column
directly, we back it out as the complement of DTC: `1 - d2c_sales_pct`. 
Two observations have `d2c_sales_pct`>100 (James City, VA and Rockdale, GA) and are assigned NA.
```{r}
dat <- dat %>%
  mutate(
    intermediated_sales_pct_derived = if_else(
      d2c_sales_pct > 100,
      NA,
      100 - d2c_sales_pct
    )
  )
```

### Step 3 — compute the SPMA index
```{r}
spma_index <- dat %>%
  mutate(
    
    has_hub  = !is.na(number_foodhub) & number_foodhub > 0,
    dtc_sales          = all_sales * d2c_sales_pct/100,
    interm_sales       = all_sales * intermediated_sales_pct_derived/100,
    SPMA = dtc_sales / (distance_nearest_farmersmarket_mile^2) +
           interm_sales  / (distance_nearest_hub_mile^2)
  )

# quick look
spma_index %>%
  summarise(n = n(),
            n_with_hub = sum(has_hub, na.rm = TRUE),
            spma_na = sum(is.na(SPMA)),
            spma_min = min(SPMA, na.rm = TRUE),
            spma_median = median(SPMA, na.rm = TRUE),
            spma_max = max(SPMA, na.rm = TRUE))
```

### Step 4 — Log & Rescale the index to 0-100
```{r}
rescale_0_100 <- function(x) {
  (x - min(x, na.rm = TRUE)) / (max(x, na.rm = TRUE) - min(x, na.rm = TRUE)) * 100
}

spma_index$SPMA_log   <- log(spma_index$SPMA + 1)   # avoid log(0)
spma_index$SPMA_0_100 <- rescale_0_100(spma_index$SPMA_log)

# clip point for the color scale (95th pct); reported in the caption
clip_hi <- as.numeric(quantile(spma_index$SPMA_0_100, 0.95, na.rm = TRUE))
n_na    <- sum(is.na(spma_index$SPMA_0_100))
```

save the index 
```{r}
spma_index %>%
  select(county_fips, state, county_name = county_name_hub,
         distance_nearest_farmersmarket_mile, distance_nearest_hub_mile,
         number_foodhub, has_hub, all_sales,
         d2c_sales_pct, local_sales_pct, intermediated_sales_pct_derived,
         SPMA, SPMA_0_100,share) %>%
  write.csv("../data/outcome/spma_index.csv", row.names = FALSE)
```
