---
title: "an_01a_spma_index"
author: "Yuanyuan Wen, Pragati Dahal"
date: "2026-07-09"
output: html_document
---
## Purpose

Compute the Small-Producer Market Access (SPMA) index for every county from
`distance_SPMA_merged.csv`, then map it as a US county choropleth.

## The SPMA index

For county \(i\), let

- \(d^{\text{FM}}_i\) = distance to the nearest **farmers market** (miles),
- \(d^{\text{hub}}_i\) = distance to the nearest **food hub** (miles),

and define the three sales pieces from the county's total sales `all_sales`:

$$
\text{DTC}_i = \text{all\_sales}_i \times \text{d2c\_sales\_pct}_i,\qquad
\text{LocalInt}_i = \text{all\_sales}_i \times (\text{local\_sales\_pct}_i - \text{d2c\_sales\_pct}_i),\qquad
\text{Int}_i = \text{all\_sales}_i \times \text{intermediated\_sales\_pct}_i .
$$

Then

$$
\text{SPMA}_i =
\begin{cases}
\dfrac{\text{DTC}_i}{\left(d^{\text{FM}}_i\right)^2} \;+\; \dfrac{\text{LocalInt}_i}{\left(d^{\text{hub}}_i\right)^2},
  & \text{if county } i \text{ has a food hub,}\\[2.2ex]
\dfrac{\text{DTC}_i}{\left(d^{\text{FM}}_i\right)^2} \;+\; \dfrac{\text{Int}_i}{\left(d^{\text{hub}}_i\right)^2},
  & \text{if county } i \text{ has no food hub.}
\end{cases}
$$

A county "has a food hub" when `number_foodhub > 0`. Note that when a county has **no** food
hub its nearest food hub is, by construction, **outside** the county, and its nearest farmers
market distance is measured from the county centroid — so `distance_nearest_hub_mile` and
`distance_nearest_farmersmarket_mile` already carry the "within/outside the county" meaning.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Load the library
```{r}
library(readr)
library(dplyr)
library(sf)
library(tigris)
library(ggplot2)
options(tigris_use_cache = TRUE)
```

### Step 1 — read the merged county table
```{r}
dat <- read_csv("../data/outcome/distance_SPMA_merged.csv") %>%
  mutate(county_fips = sprintf("%05d", as.integer(county_fips)))
```

### Step 2 — total sales (`all_sales`)
`all_sales` (total county sales, in dollars) is a multiplicative factor on the whole index.
Until it is added to `SPMA_index_variables.csv`, we fall back to `all_sales = 1` so the code
still runs; then the index reflects the sales **shares** only. Once the real column exists it
is used automatically.
```{r}
if (!"all_sales" %in% names(dat)) {
  message("`all_sales` not found -> using all_sales = 1 (share-based index placeholder).")
  dat$all_sales <- 1
}
```

### Step 2b — derive intermediated sales share from DTC share
`d2c_sales_pct` and `all_sales` have the fewest missing values of any of the sales
variables, so rather than relying on the sparser `intermediated_sales_pct` column
directly, we back it out as the complement of DTC: `1 - d2c_sales_pct`. Where
`d2c_sales_pct` exceeds 1 (a data quality issue — a sales share can't exceed 100%
by definition, see earlier note on Census of Ag disclosure artifacts), the derived
intermediated share is set to 0 rather than a negative number.
```{r}
dat <- dat %>%
  mutate(
    intermediated_sales_pct_derived = if_else(
      d2c_sales_pct > 1,
      0,
      1 - d2c_sales_pct
    )
  )
```

### Step 3 — compute the SPMA index
```{r}
spma_index <- dat %>%
  mutate(
    has_hub  = !is.na(number_foodhub) & number_foodhub > 0,

    dtc_sales          = all_sales * d2c_sales_pct,
    interm_sales       = all_sales * (1 - d2c_sales_pct),
    
    SPMA = dtc_sales / (distance_nearest_farmersmarket_mile^2) +
           interm_sales  / (distance_nearest_hub_mile^2)
  )

# quick look
spma_index %>%
  summarise(n = n(),
            n_with_hub = sum(has_hub, na.rm = TRUE),
            spma_na = sum(is.na(SPMA)),
            spma_min = min(SPMA, na.rm = TRUE),
            spma_median = median(SPMA, na.rm = TRUE),
            spma_max = max(SPMA, na.rm = TRUE))
```

### Step 4 — US county choropleth
```{r}
# lower-48 + DC county polygons (drop AK/HI/territories)
counties_sf <- counties(cb = TRUE, year = 2023, progress_bar = FALSE) %>%
  filter(!STATEFP %in% c("02", "15", "60", "66", "69", "72", "78")) %>%
  transmute(county_fips = GEOID) %>%
  st_transform(5070)

states_sf <- states(cb = TRUE, year = 2023, progress_bar = FALSE) %>%
  filter(!STATEFP %in% c("02", "15", "60", "66", "69", "72", "78")) %>%
  st_transform(5070)

map_df <- counties_sf %>% left_join(spma_index, by = "county_fips")

p <- ggplot(map_df) +
  geom_sf(aes(fill = SPMA), color = NA) +
  geom_sf(data = states_sf, fill = NA, color = "grey30", linewidth = 0.2) +
  scale_fill_viridis_c(option = "magma", direction = -1, trans = "log10",
                       labels = scales::label_number(), na.value = "grey88",
                       name = "SPMA\n(log scale)") +
  labs(title = "Small-Producer Market Access (SPMA) index by county",
       subtitle = "Higher = better weighted access to farmers markets & food hubs",
       caption = "SPMA = DTC/d_FM^2 + (local-intermediated or intermediated)/d_hub^2") +
  theme_void(base_size = 12) +
  theme(plot.background  = element_rect(fill = "white", color = NA),
        panel.background = element_rect(fill = "white", color = NA),
        plot.title    = element_text(face = "bold"),
        plot.subtitle = element_text(color = "grey35"),
        plot.caption  = element_text(color = "grey55"),
        legend.position = c(0.92, 0.25))
p
```

### Step 5 — save
```{r}
dir.create("../figures", showWarnings = FALSE)
ggsave("../figures/spma_choropleth.png", p, width = 11, height = 7, dpi = 150)

spma_index %>%
  select(county_fips, state, county_name = county_name_hub,
         distance_nearest_farmersmarket_mile, distance_nearest_hub_mile,
         number_foodhub, has_hub, all_sales,
         d2c_sales_pct, local_sales_pct, intermediated_sales_pct_derived, SPMA) %>%
  write.csv("../data/outcome/spma_index.csv", row.names = FALSE)
```
