---
title: "an_02_spma_simulation"
author: "Yuanyuan Wen, Pragati Dahal, Xiaoyi Zhao"
date: "2026-07-16"
output: html_document
---
## Purpose

**Goal:** Simulate counterfactual SPMA scenarios for two paired counties:

1. If a county did not have its nearest food hub, what would its SPMA be?
2. If a county had access to its paired county's nearest food hub, what would its SPMA be?

**Steps:**

1. Identify the distance from each county centroid to its second nearest food hub.
2. Assign one county's nearest food hub to its paired county.

**Pair:** Yakima County, WA and Okanogan County, WA. Change lines 38-39 to change a county pair.

**Data:** County centroids are taken from `distance_SPMA_merged.csv`; food hub locations come from `foodhub_2026-75161420.xlsx`.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the library
```{r}
library(sf)        # the workhorse: stores and manipulates spatial data
library(dplyr)     # regular data wrangling once things look like data frames
library(readr)     # read the merged csv that carries the county centroids
library(readxl)
library(units)
```

### Load data
Get the county centroids for the two target counties
```{r}
target_fips <- c("53077",   # Yakima County, WA
                 "53047")   # Okanogan County, WA

county_centroid <- read_csv("../data/outcome/distance_SPMA_merged.csv") %>%
  mutate(county_fips = sprintf("%05d", as.integer(county_fips))) %>%
  filter(county_fips %in% target_fips) %>%
  transmute(county_fips, state, county_name = county_name_hub,
            centroid_x, centroid_y,
            all_sales, d2c_sales_pct,distance_nearest_farmersmarket_mile, share) %>%
  st_as_sf(coords = c("centroid_x", "centroid_y"), crs = 5070)  # 5070 = NAD83 Albers, in metres

```

Load the food hub data
```{r}
hub_raw <- read_excel("../data/source/foodhub_2026-75161420.xlsx", col_types = "text")

hub_directory <- hub_raw %>%
  select(listing_id, listing_name, location_x, location_y) %>%
  rename(lon = location_x, lat = location_y) %>%
  mutate(lon = as.numeric(lon),
         lat = as.numeric(lat)) %>%
  filter(!is.na(lon), !is.na(lat))
```

make the hubs spatial, then convert them into meters by 5070 so that we can get the distance from the centroid to the hub
```{r}
hubs_sf  <- st_as_sf(hub_directory, coords = c("lon", "lat"), crs = 4326)
hub_proj <- st_transform(hubs_sf, 5070)

#check whether two are in the same scale
st_crs(hub_proj) == st_crs(county_centroid)
```

identify the distance between the centroid and every hub
```{r}
distance_hub <- st_distance(county_centroid, hub_proj)

# (number of target counties) x (number of food hubs)
dim(distance_hub)
```

for each of the row in the distance matrix, rank the hubs by distance:
```{r}
first_idx   <- apply(distance_hub, 1, function(d) order(d)[1])
second_idx  <- apply(distance_hub, 1, function(d) order(d)[2])

first_dist  <- apply(distance_hub, 1, function(d) sort(d)[1])
second_dist <- apply(distance_hub, 1, function(d) sort(d)[2])
```

for each county, report the nearest and the second nearest food hub
```{r}
second_hub_distance <- county_centroid %>%
  st_drop_geometry() %>%
  mutate(
    nearest_hub_name                    = hub_directory$listing_name[first_idx],
    nearest_hub_id                      = hub_directory$listing_id[first_idx],
    distance_nearest_hub_mile           = set_units(set_units(first_dist, "m"), "mi") %>% as.numeric(),

    second_hub_name                     = hub_directory$listing_name[second_idx],
    second_hub_id                       = hub_directory$listing_id[second_idx],
    distance_2rd_nearest_hub_mile = set_units(set_units(second_dist, "m"), "mi") %>% as.numeric()
  )

second_hub_distance
```


## Calculate their new SPMA
```{r}
spma_index <- second_hub_distance %>%
  mutate(
    dtc_sales    = all_sales * d2c_sales_pct/100,
    interm_sales = all_sales * (100 - d2c_sales_pct)/100,
    SPMA = dtc_sales / (distance_nearest_farmersmarket_mile^2) +
           interm_sales  / (distance_nearest_hub_mile^2),
    
    SPMA_2rd = dtc_sales / (distance_nearest_farmersmarket_mile^2) +
               interm_sales / (distance_2rd_nearest_hub_mile^2),
    
    SPMA_exchange = dtc_sales / (distance_nearest_farmersmarket_mile^2) +
                    interm_sales / (rev(distance_nearest_hub_mile)^2)
  )

spma_index

```

rescale SPMA
```{r}
temp_dat <- read_csv("../data/outcome/SPMA_index.csv")
min <- min(log(temp_dat$SPMA+1), na.rm = TRUE)
max <- max(log(temp_dat$SPMA+1), na.rm = TRUE)

rescale_0_100 <- function(x) {
  (x - min) / (max - min) * 100
}

spma_index$SPMA_2rd_log <- log(spma_index$SPMA_2rd + 1)   # avoid log(0)
spma_index$SPMA_exchange_log <- log(spma_index$SPMA_exchange + 1)   # avoid log(0)
spma_index$SPMA_2rd_0_100 <- rescale_0_100(spma_index$SPMA_2rd_log)
spma_index$SPMA_exchange_0_100 <- rescale_0_100(spma_index$SPMA_exchange_log)
spma_index$SPMA_log <- log(spma_index$SPMA + 1)   # avoid log(0)
spma_index$SPMA_0_100 <- rescale_0_100(spma_index$SPMA_log)

spma_index %>% select(county_name, SPMA_0_100, SPMA_2rd_0_100,SPMA_exchange_0_100)

```
## Save data for the pair
```{r}
write.csv(spma_index,
          "../data/outcome/spma_index_yakima_okanogan.csv",
          row.names = FALSE)
```

# another pair of target fips : 

### Load data
Get the county centroids for the two target counties
```{r}
target_fips <- c("45045",     #Greenville County, SC
                 "13177")   #Lee County, GA

county_centroid <- read_csv("../data/outcome/distance_SPMA_merged.csv") %>%
  mutate(county_fips = sprintf("%05d", as.integer(county_fips))) %>%
  filter(county_fips %in% target_fips) %>%
  transmute(county_fips, state, county_name = county_name_hub,
            centroid_x, centroid_y,
            all_sales, d2c_sales_pct,distance_nearest_farmersmarket_mile, share) %>%
  st_as_sf(coords = c("centroid_x", "centroid_y"), crs = 5070)  # 5070 = NAD83 Albers, in metres


```

Load the food hub data
```{r}
hub_raw <- read_excel("../data/source/foodhub_2026-75161420.xlsx", col_types = "text")

hub_directory <- hub_raw %>%
  select(listing_id, listing_name, location_x, location_y) %>%
  rename(lon = location_x, lat = location_y) %>%
  mutate(lon = as.numeric(lon),
         lat = as.numeric(lat)) %>%
  filter(!is.na(lon), !is.na(lat))
```

make the hubs spatial, then convert them into meters by 5070 so that we can get the distance from the centroid to the hub
```{r}
hubs_sf  <- st_as_sf(hub_directory, coords = c("lon", "lat"), crs = 4326)
hub_proj <- st_transform(hubs_sf, 5070)

#check whether two are in the same scale
st_crs(hub_proj) == st_crs(county_centroid)
```

identify the distance between the centroid and every hub
```{r}
distance_hub <- st_distance(county_centroid, hub_proj)

# (number of target counties) x (number of food hubs)
dim(distance_hub)
```

for each of the row in the distance matrix, rank the hubs by distance:
```{r}
first_idx   <- apply(distance_hub, 1, function(d) order(d)[1])
second_idx  <- apply(distance_hub, 1, function(d) order(d)[2])

first_dist  <- apply(distance_hub, 1, function(d) sort(d)[1])
second_dist <- apply(distance_hub, 1, function(d) sort(d)[2])
```

for each county, report the nearest and the second nearest food hub
```{r}
second_hub_distance <- county_centroid %>%
  st_drop_geometry() %>%
  mutate(
    nearest_hub_name                    = hub_directory$listing_name[first_idx],
    nearest_hub_id                      = hub_directory$listing_id[first_idx],
    distance_nearest_hub_mile           = set_units(set_units(first_dist, "m"), "mi") %>% as.numeric(),

    second_hub_name                     = hub_directory$listing_name[second_idx],
    second_hub_id                       = hub_directory$listing_id[second_idx],
    distance_2rd_nearest_hub_mile = set_units(set_units(second_dist, "m"), "mi") %>% as.numeric()
  )

second_hub_distance
```


## Calculate their new SPMA
```{r}
spma_index <- second_hub_distance %>%
  mutate(
    dtc_sales    = all_sales * d2c_sales_pct/100,
    interm_sales = all_sales * (100 - d2c_sales_pct)/100,
    SPMA = dtc_sales / (distance_nearest_farmersmarket_mile^2) +
           interm_sales  / (distance_nearest_hub_mile^2),
    
    SPMA_2rd = dtc_sales / (distance_nearest_farmersmarket_mile^2) +
               interm_sales / (distance_2rd_nearest_hub_mile^2),
    
    SPMA_exchange = dtc_sales / (distance_nearest_farmersmarket_mile^2) +
                    interm_sales / (rev(distance_nearest_hub_mile)^2)
  )

spma_index

```

rescale SPMA
```{r}
temp_dat <- read_csv("../data/outcome/SPMA_index.csv")
min <- min(log(temp_dat$SPMA+1), na.rm = TRUE)
max <- max(log(temp_dat$SPMA+1), na.rm = TRUE)

rescale_0_100 <- function(x) {
  (x - min) / (max - min) * 100
}

spma_index$SPMA_2rd_log <- log(spma_index$SPMA_2rd + 1)   # avoid log(0)
spma_index$SPMA_exchange_log <- log(spma_index$SPMA_exchange + 1)   # avoid log(0)
spma_index$SPMA_2rd_0_100 <- rescale_0_100(spma_index$SPMA_2rd_log)
spma_index$SPMA_exchange_0_100 <- rescale_0_100(spma_index$SPMA_exchange_log)
spma_index$SPMA_log <- log(spma_index$SPMA + 1)   # avoid log(0)
spma_index$SPMA_0_100 <- rescale_0_100(spma_index$SPMA_log)

spma_index %>% select(county_name, SPMA_0_100, SPMA_2rd_0_100,SPMA_exchange_0_100)

```
## Save data for the pair
```{r}
write.csv(spma_index,
          "../data/outcome/spma_index_greenville_lee.csv",
          row.names = FALSE)
```
