---
title: "an_01b_spma_index"
author: "Xiaoyi Zhao"
date: "2026-07-15"
output: html_document
---
## Purpose

Add the more outstanding hub plot and the 3D effect on the US county choropleth.

## The SPMA index

For county \(i\), let

- \(d^{\text{FM}}_i\) = distance from the county centroid to the nearest **farmers market** (miles),
- \(d^{\text{hub}}_i\) = distance from the county centroid to the nearest **food hub** (miles),

and define the three sales pieces from the county's total sales `all_sales`:

$$
\text{DTC}_i = \text{all_sales}_i \times \text{d2c_sales_pct}_i/100,\qquad
\text{Int}_i = \text{all_sales}_i \times \text{intermediated_sales_pct}_i/100 .
$$

Then

$$
\text{SPMA}_i =
\dfrac{\text{DTC}_i}{\left(d^{\text{FM}}_i\right)^2} \;+\; \dfrac{\text{Int}_i}{\left(d^{\text{hub}}_i\right)^2}
$$

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Load the library
```{r}
library(readr)
library(dplyr)
library(sf)
library(tigris)
library(ggplot2)
library(readxl)
library(ggnewscale)
library(ggtext) # adjust the text style in the ggplot

# To install rayshader
# link to the package https://www.rayshader.com/
library(rayshader)
```

### Step 1 — read the merged county table
```{r}
dat <- read_csv("../data/outcome/distance_SPMA_merged.csv") %>%
  mutate(county_fips = sprintf("%05d", as.integer(county_fips)))
```


### Step 2 — derive intermediated sales share from DTC share
`d2c_sales_pct` and `all_sales` have the fewest missing values of any of the sales
variables, so rather than relying on the sparser `intermediated_sales_pct` column
directly, we back it out as the complement of DTC: `1 - d2c_sales_pct`. 
Two observations have `d2c_sales_pct`>100 (James City, VA and Rockdale, GA) and are assigned NA.
```{r}
dat <- dat %>%
  mutate(
    intermediated_sales_pct_derived = if_else(
      d2c_sales_pct > 100,
      NA,
      100 - d2c_sales_pct
    )
  )
```

### Step 3 — compute the SPMA index
```{r}
spma_index <- dat %>%
  mutate(
    
    has_hub  = !is.na(number_foodhub) & number_foodhub > 0,
    dtc_sales          = all_sales * d2c_sales_pct/100,
    interm_sales       = all_sales * intermediated_sales_pct_derived/100,
    SPMA = dtc_sales / (distance_nearest_farmersmarket_mile^2) +
           interm_sales  / (distance_nearest_hub_mile^2)
  )

# quick look
spma_index %>%
  summarise(n = n(),
            n_with_hub = sum(has_hub, na.rm = TRUE),
            spma_na = sum(is.na(SPMA)),
            spma_min = min(SPMA, na.rm = TRUE),
            spma_median = median(SPMA, na.rm = TRUE),
            spma_max = max(SPMA, na.rm = TRUE))
```

### Step 4 — Log & Rescale the index to 0-100
```{r}
rescale_0_100 <- function(x) {
  (x - min(x, na.rm = TRUE)) / (max(x, na.rm = TRUE) - min(x, na.rm = TRUE)) * 100
}

spma_index$SPMA_log   <- log(spma_index$SPMA + 1)   # avoid log(0)
spma_index$SPMA_0_100 <- rescale_0_100(spma_index$SPMA_log)

# clip point for the color scale (95th pct); reported in the caption
clip_hi <- as.numeric(quantile(spma_index$SPMA_0_100, 0.95, na.rm = TRUE))
n_na    <- sum(is.na(spma_index$SPMA_0_100))
```

### Step 4b - Food hub location
```{r}
drop_fp <- c("02", "15", "60", "66", "69", "72", "78")

foodhub_raw <- read_excel("../data/source/foodhub_2026-75161420.xlsx")

foodhub_directory <- foodhub_raw %>%
  select(listing_id, listing_name, location_x, location_y, products) %>%
  rename(lon = location_x, lat = location_y) %>%
  mutate(lon = as.numeric(lon),
         lat = as.numeric(lat)) %>%
  filter(!is.na(lon), !is.na(lat))

hubs_sf <- st_as_sf(foodhub_directory, coords = c("lon", "lat"), crs = 4326)

county_lookup <- counties(cb = TRUE, year = 2023, progress_bar = FALSE) %>%
  select(STATEFP, GEOID, county_name = NAME) %>%
  st_transform(4326)

hubs_pts <- st_join(hubs_sf, county_lookup, join = st_within) %>%
  filter(!STATEFP %in% drop_fp) %>%
  st_transform(5070)

crop_lu <- tibble::tribble(
  ~NAME,            ~icon,
  "Washington",     "🍎",
  "Oregon",         "🍐",
  "California",     "🍇",
  "Wisconsin",      "🫐",
  "Michigan",       "🍒",
  "Maine",          "🫐",
  "Florida",        "🍊",
  "Georgia",        "🍑",
  "South Carolina", "🍑",
  "Texas",          "🍈"
)
```
### Step 5 — US county choropleth
```{r}
# lower-48 + DC county polygons (drop AK/HI/territories)
counties_sf <- counties(cb = TRUE, year = 2023, progress_bar = FALSE) %>%
  filter(!STATEFP %in% drop_fp) %>%
  transmute(county_fips = GEOID) %>%
  st_transform(5070)

states_sf <- states(cb = TRUE, year = 2023, progress_bar = FALSE) %>%
  filter(!STATEFP %in% drop_fp) %>%
  select(NAME) %>%
  st_transform(5070)

crop_pts <- states_sf %>%
  inner_join(crop_lu, by = "NAME") %>%
  st_point_on_surface()

map_df <- counties_sf %>%
  left_join(spma_index, by = "county_fips")

p <- ggplot(map_df) +
  geom_sf(aes(fill = SPMA_0_100), color = NA) +
  geom_sf(data = states_sf, fill = NA, color = "grey30", linewidth = 0.2) +
  geom_sf(data = hubs_pts, shape = 21, fill = "yellow",
          color = "grey50", size = 1.2, stroke = 0.25, alpha = 0.5) +
  scale_fill_distiller(
    palette   = "Blues",
    direction = 1,
    limits    = c(0, clip_hi),
    oob       = scales::squish,
    labels    = scales::label_number(accuracy = 1),
    na.value  = "grey90",
    name      = "SPMA\n(0-100)"
  ) +
  coord_sf(crs = 5070) +
  geom_sf_text(data = crop_pts, aes(label = icon), size = 7) +
  labs(
    title    = "Small-Producer Market Access (SPMA) index by county",
    subtitle = "Higher = better weighted access to farmers markets & food hubs",
    caption = paste0(
  "Grey polygons = no data (", n_na, " counties). Color scale clipped at ",
  round(clip_hi), "; counties above are shown at the darkest shade.\n",
  "Yellow dots = food hub locations. Fruit icons mark leading fruit/nut states ",
  "(USDA ERS, Fruit and Tree Nuts Outlook, FTS-384, March 2026); illustrative, not to scale.")
  ) +
  theme_void(base_size = 12) +
  theme(
    plot.background  = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    plot.title    = element_text(face = "bold"),
    plot.subtitle = element_text(color = "grey35"),
    plot.caption  = element_text(size = 8, color = "grey55", hjust = 0),
    legend.position = c(0.92, 0.25)
  )

# 1. we set the minimum to be floor of the 3D map
floor_val <- min(map_df$SPMA_0_100, na.rm = TRUE)

# 2. this makes the height match the color clipping
map_df$SPMA_plot <- pmin(map_df$SPMA_0_100, clip_hi)

p3d <- p +
  scale_fill_gradientn(
  colors = c("white", "#c6dbef", "#4292c6", "#08306b"),
  values = scales::rescale(c(0, 0.35, 0.7, 1)),
  limits = c(floor_val, clip_hi),
  oob    = scales::squish,
  labels = scales::label_number(accuracy = 1),
  na.value = "grey90",
  name   = "SPMA\n(0-100)"
)+
  theme(
    legend.position = "right",         
    plot.caption = element_blank()     
  )

#Takes a ggplot2 object (or a list of two ggplot2 objects) and uses the fill or color aesthetic to transform the plot into a 3D surface. 
plot_gg(
  p3d,
  multicore  = TRUE,
  width      = 9,  height = 5.5,   # match the wide US-map aspect ratio
  scale      = 80,                # vertical exaggeration — tune 150–350
  height_aes = "fill",             # extrude from the fill variable
  raytrace   = TRUE,
  windowsize = c(1400, 900),
  zoom       = 0.7,
  phi        = 40,                 # camera elevation
  theta      = 0,                  # facing north-up
  sunangle   = 315,
  shadow_intensity = 0.5
)


# After plot_gg() has the scene open, adjust the camera without re-rendering
# This will give us a static plot
render_camera(
  phi   = 65,     # 90 = straight down, 40 = your current dramatic tilt
  theta = 0,      # north stays up, no rotation
  zoom  = 0.55,
  fov   = 0       # 0 = orthographic — no perspective distortion, best for papers
)

# make the snapshot of the plot
render_snapshot("spma_3d_static.png",
                software_render = TRUE,
                width = 3000, height = 1900)
```

### Step 6 — save
```{r}
dir.create("../figures", showWarnings = FALSE)
ggsave("../figures/spma_choropleth_overlay.png", p, width = 11, height = 7, dpi = 150,
       device = ragg::agg_png)
spma_index %>%
  select(county_fips, state, county_name = county_name_hub,
         distance_nearest_farmersmarket_mile, distance_nearest_hub_mile,
         number_foodhub, has_hub, all_sales,
         d2c_sales_pct, local_sales_pct, intermediated_sales_pct_derived,
         SPMA, SPMA_0_100) %>%
  write.csv("../data/outcome/spma_index.csv", row.names = FALSE)
```
