---
title: "an_01b_spma_index"
author: "Xiaoyi Zhao"
date: "2026-07-15"
output: html_document
---
## Purpose

Add the more outstanding hub plot and the 3D effect on the US county choropleth.

## The SPMA index

For county \(i\), let

- \(d^{\text{FM}}_i\) = distance from the county centroid to the nearest **farmers market** (miles),
- \(d^{\text{hub}}_i\) = distance from the county centroid to the nearest **food hub** (miles),

and define the three sales pieces from the county's total sales `all_sales`:

$$
\text{DTC}_i = \text{all_sales}_i \times \text{d2c_sales_pct}_i/100,\qquad
\text{Int}_i = \text{all_sales}_i \times \text{intermediated_sales_pct}_i/100 .
$$

Then

$$
\text{SPMA}_i =
\dfrac{\text{DTC}_i}{\left(d^{\text{FM}}_i\right)^2} \;+\; \dfrac{\text{Int}_i}{\left(d^{\text{hub}}_i\right)^2}
$$

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Load the library
```{r}
# to install the packages

# install.packages(packages <- c(
#   "sf",
#   "dplyr",
#   "ggplot2",
#   "tidyr",
#   "purrr",
#   "ggplot2",
#   "scales",
#   "tidycensus"
# )
# )
# 
# install.packages(c("showtext", "sysfonts"))

install.packages("rsvg")
library(rsvg)

library(showtext)
library(sysfonts)

install.packages("ggimage")
library(ggimage)


library(showtext)
library(sysfonts)

```

```{r}
library(sf)
library(dplyr)
library(tidyr)
library(purrr)
library(scales)
library(terra)

library(tidycensus)
library(magick)
library(readr)
library(sf)
library(tigris)
library(readxl)
library(ggnewscale)
library(ggtext) # adjust the text style in the ggplot

library(showtext)
library(sysfonts)

library(ggplot2)
library(patchwork)

```

### Step 1 — read the merged county table
```{r}
dat <- read_csv("../data/outcome/distance_SPMA_merged.csv") %>%
  mutate(county_fips = sprintf("%05d", as.integer(county_fips)))
```


### Step 2 — derive intermediated sales share from DTC share
`d2c_sales_pct` and `all_sales` have the fewest missing values of any of the sales
variables, so rather than relying on the sparser `intermediated_sales_pct` column
directly, we back it out as the complement of DTC: `1 - d2c_sales_pct`. 
Two observations have `d2c_sales_pct`>100 (James City, VA and Rockdale, GA) and are assigned NA.
```{r}
dat <- dat %>%
  mutate(
    intermediated_sales_pct_derived = if_else(
      d2c_sales_pct > 100,
      NA,
      100 - d2c_sales_pct
    )
  )
```

### Step 3 — compute the SPMA index
```{r}
spma_index <- dat %>%
  mutate(
    
    has_hub  = !is.na(number_foodhub) & number_foodhub > 0,
    dtc_sales          = all_sales * d2c_sales_pct/100,
    interm_sales       = all_sales * intermediated_sales_pct_derived/100,
    SPMA = dtc_sales / (distance_nearest_farmersmarket_mile^2) +
           interm_sales  / (distance_nearest_hub_mile^2)
  )

# quick look
spma_index %>%
  summarise(n = n(),
            n_with_hub = sum(has_hub, na.rm = TRUE),
            spma_na = sum(is.na(SPMA)),
            spma_min = min(SPMA, na.rm = TRUE),
            spma_median = median(SPMA, na.rm = TRUE),
            spma_max = max(SPMA, na.rm = TRUE))
```

### Step 4 — Log & Rescale the index to 0-100
```{r}
rescale_0_100 <- function(x) {
  (x - min(x, na.rm = TRUE)) / (max(x, na.rm = TRUE) - min(x, na.rm = TRUE)) * 100
}

spma_index$SPMA_log   <- log(spma_index$SPMA + 1)   # avoid log(0)
spma_index$SPMA_0_100 <- rescale_0_100(spma_index$SPMA_log)

# clip point for the color scale (95th pct); reported in the caption
clip_hi <- as.numeric(quantile(spma_index$SPMA_0_100, 0.95, na.rm = TRUE))
n_na    <- sum(is.na(spma_index$SPMA_0_100))
```

### Step 4b - Food hub location
```{r}
drop_fp <- c("02", "15", "60", "66", "69", "72", "78")

foodhub_raw <- read_excel("../data/source/foodhub_2026-75161420.xlsx")

foodhub_directory <- foodhub_raw %>%
  select(listing_id, listing_name, location_x, location_y, products) %>%
  rename(lon = location_x, lat = location_y) %>%
  mutate(lon = as.numeric(lon),
         lat = as.numeric(lat)) %>%
  filter(!is.na(lon), !is.na(lat))

hubs_sf <- st_as_sf(foodhub_directory, coords = c("lon", "lat"), crs = 4326)

county_lookup <- counties(cb = TRUE, year = 2023, progress_bar = FALSE) %>%
  select(STATEFP, GEOID, county_name = NAME) %>%
  st_transform(4326)

hubs_pts <- st_join(hubs_sf, county_lookup, join = st_within) %>%
  filter(!STATEFP %in% drop_fp) %>%
  st_transform(5070)

crop_lu <- tibble::tribble(
  ~NAME,            ~icon,
  "Washington",     "🍎",
  "Oregon",         "🍐",
  "California",     "🍇",
  "Wisconsin",      "🫐",
  "Michigan",       "🍒",
  "Maine",          "🫐",
  "Florida",        "🍊",
  "Georgia",        "🍑",
  "South Carolina", "🍑",
  "Texas",          "🍈"
)
```

```{r}
svg_files <- list.files(
  "../assets/svg",
  pattern = "\\.svg$",
  full.names = TRUE
)

dir.create("../assets/png", showWarnings = FALSE, recursive = TRUE)

for (svg_file in svg_files) {

  png_file <- file.path(
    "../assets/png",
    paste0(tools::file_path_sans_ext(basename(svg_file)), ".png")
  )

  rsvg_png(
    svg = svg_file,
    file = png_file,
    width = 1024,
    height = 1024
  )
}
```

### Step 5 — US county choropleth
```{r}
# lower-48 + DC county polygons (drop AK/HI/territories)
counties_sf <- counties(cb = TRUE, year = 2023, progress_bar = FALSE) %>%
  filter(!STATEFP %in% drop_fp) %>%
  transmute(county_fips = GEOID) %>%
  st_transform(5070)

states_sf <- states(cb = TRUE, year = 2023, progress_bar = FALSE) %>%
  filter(!STATEFP %in% drop_fp) %>%
  select(NAME) %>%
  st_transform(5070)

crop_pts <- states_sf %>%
  inner_join(crop_lu, by = "NAME") %>%
  st_point_on_surface()

map_df <- counties_sf %>%
  left_join(spma_index, by = "county_fips")

crop_pts <- crop_pts %>% 
  mutate(
    image = case_when(
      NAME == "California"     ~ "../assets/png/grapes.png",
      NAME == "Georgia"        ~ "../assets/png/peach.png",
      NAME == "South Carolina" ~ "../assets/png/peach.png",
      NAME == "Michigan"       ~ "../assets/png/cherry.png",
      NAME == "Florida"        ~ "../assets/png/orange.png",
      NAME == "Maine"          ~ "../assets/png/blueberry.png",
      NAME == "Washington"     ~ "../assets/png/apple.png",
      NAME == "Texas"          ~ "../assets/png/melon.png",
      NAME == "Wisconsin"      ~ "../assets/png/blueberry.png",
      NAME == "Oregon"         ~ "../assets/png/pear.png",
      TRUE                     ~ NA_character_
    )
  )

crop_icon_df <- crop_pts %>% 
  mutate(
    x = sf::st_coordinates(.)[, 1],
    y = sf::st_coordinates(.)[, 2]
  ) %>% 
  sf::st_drop_geometry()
```


```{r, fig.width= 12, fig.height= 5}

font_add_google(
  name = "DM Serif Display",
  family = "title_font"
)

font_add_google(
  name = "DM Sans",
  family = "body_font"
)

showtext::showtext_auto(TRUE)

fig_01c_annotated <- ggplot(map_df) +
  geom_sf(aes(fill = SPMA_0_100), color = NA) + 
  geom_sf(data = states_sf, fill = NA, color = "grey55", linewidth = 0.25) + 
  geom_sf(data = hubs_pts, shape = 21, fill = "#FFD23F",
          color = "grey50", size = 1.2, stroke = 0.25, alpha = 0.5) +
  scale_fill_gradientn(
    colours =  c(
  "#EFF6FF",
  "#BFDBFE",
  "#60A5FA",
  "#2563EB",
  "#1E3A8A"
),
    limits = c(0, 100),
    oob = scales::squish,
    na.value = "#F0F0EC",
    name = NULL
  )  + 
  coord_sf(crs = 5070) +
  ggimage::geom_image(
  data = crop_icon_df,
  aes(
    x = x,
    y = y,
    image = image
  ),
  size = 0.09,
  asp = 1,
  inherit.aes = FALSE) +
  theme_void(base_family = "body_font") +
  labs(
    title = "Small-Producers' Market Accessibility(SPMA)",
    subtitle = paste0(
      "How market access connects fruit and tree nut producers ",
      "to consumers and food hubs"
    )
  ) +
  theme(
    plot.background = element_rect(
      fill = "#FAF8F2",
      color = NA
    ),
    panel.background = element_rect(
      fill = "#FAF8F2",
      color = NA
    ),

    plot.title = element_text(
      family = "title_font",
      size = 25,
      lineheight = 0.88,
      color = "#123B65",
      face = "plain",
      hjust = 0,
      margin = margin(b = 1)
    ),

    plot.subtitle = element_text(
      family = "body_font",
      size = 12,
      lineheight = 1.2,
      color = "#4B5D6D",
      hjust = 0,
      margin = margin(b = 10)
    ),

    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.width = unit(2.5, "cm"),
    legend.key.height = unit(0.28, "cm"),

    plot.margin = margin(
  t = 10,
  r = 10,
  b = 10,
  l = 10
)
  )


ggsave("../figures/fig_01c.pdf", plot = fig_01c_annotated, width = 12, height = 5)
fig_01c_annotated


```

To add the note panel 

```{r}
library(ggtext)

notes_panel <- ggplot() +
  annotate(
    "text",
    x = 0,
    y = 0.92,
    label = "Food hubs expand market access",
    hjust = 0,
    vjust = 1,
    family = "body_font",
    fontface = "bold",
    size = 5.2,
    color = "#123B65"
  ) +
  annotate(
    "text",
    x = 0,
    y = 0.82,
    label = paste0(
      "Counties closer to food hubs tend to have\n",
      "stronger access to markets and buyers.\n",
      "Yellow dots identify food hub locations."
    ),
    hjust = 0,
    vjust = 1,
    family = "body_font",
    size = 4,
    lineheight = 1.15,
    color = "#4B5D6D"
  ) +
  annotate(
    "segment",
    x = -0.06,
    xend = -0.06,
    y = 0.75,
    yend = 0.96,
    linewidth = 1.4,
    color = "#FFD23F"
  ) +
  annotate(
    "text",
    x = 0,
    y = 0.58,
    label = "Accessibility is multidimensional",
    hjust = 0,
    vjust = 1,
    family = "body_font",
    fontface = "bold",
    size = 5.2,
    color = "#123B65"
  ) +
  annotate(
    "text",
    x = 0,
    y = 0.48,
    label = paste0(
      "The SPMA combines direct-to-consumer access,\n",
      "food-hub connections, and market proximity.\n",
      "Darker blue represents higher accessibility."
    ),
    hjust = 0,
    vjust = 1,
    family = "body_font",
    size = 4,
    lineheight = 1.15,
    color = "#4B5D6D"
  ) +
  annotate(
    "segment",
    x = -0.06,
    xend = -0.06,
    y = 0.41,
    yend = 0.62,
    linewidth = 1.4,
    color = "#3182BD"
  ) +
  coord_cartesian(
    xlim = c(-0.1, 1),
    ylim = c(0, 1),
    clip = "off"
  ) +
  theme_void() +
  theme(
    plot.background = element_rect(
      fill = "#FAF8F2",
      color = NA
    )
  )

rural_note <- ggplot() +
  annotate(
    "segment",
    x = 0.02,
    xend = 0.02,
    y = 0.12,
    yend = 0.88,
    linewidth = 1.5,
    color = "#A7B6C5"
  ) +
  annotate(
    "text",
    x = 0.07,
    y = 0.84,
    label = "Why are values missing in the central U.S.?",
    hjust = 0,
    vjust = 1,
    family = "body_font",
    fontface = "bold",
    size = 4,
    color = "#123B65"
  ) +
  annotate(
    "text",
    x = 0.07,
    y = 0.62,
    label = paste0(
      "Missing SPMA values mainly reflect unavailable or suppressed sales data.\n",
      "These gaps are more common in rural counties and places with fewer fruit\n",
      "and tree nut farms."
    ),
    hjust = 0,
    vjust = 1,
    family = "body_font",
    size = 3,
    lineheight = 1.15,
    color = "#4B5D6D"
  ) +
  coord_cartesian(
    xlim = c(0, 1),
    ylim = c(0, 1),
    clip = "off"
  ) +
  theme_void() +
  theme(
    plot.background = element_rect(
      fill = "#FAF8F2",
      color = NA
    ),
    plot.margin = margin(5, 10, 5, 20)
  )

left_column <- fig_01c_annotated / rural_note +
  plot_layout(
    heights = c(4.5, 1)
  )


final_layout <- (left_column | notes_panel) +
  patchwork::plot_layout(
    ncol = 2,
    widths = c(3.2, 1.25)
  ) +
  plot_annotation(
    theme = theme(
      plot.background = element_rect(
        fill = "#FAF8F2",
        color = NA
      ),
      plot.title = element_text(
        family = "title_font",
        size = 34,
        lineheight = 0.88,
        color = "#123B65",
        hjust = 0,
        margin = margin(b = 12)
      ),
      plot.subtitle = element_text(
        family = "body_font",
        size = 14,
        color = "#4B5D6D",
        hjust = 0,
        margin = margin(b = 18)
      ),
      plot.margin = margin(
        t = 30,
        r = 35,
        b = 25,
        l = 35
      )
    )
  )

final_layout

ggsave("../figures/fig_01c.pdf", plot = final_layout, width = 18, height = 8)

```


### Step 6 — save
```{r}
dir.create("../figures", showWarnings = FALSE)
ggsave("../figures/spma_choropleth_overlay.png", p, width = 11, height = 7, dpi = 150,
       device = ragg::agg_png)
spma_index %>%
  select(county_fips, state, county_name = county_name_hub,
         distance_nearest_farmersmarket_mile, distance_nearest_hub_mile,
         number_foodhub, has_hub, all_sales,
         d2c_sales_pct, local_sales_pct, intermediated_sales_pct_derived,
         SPMA, SPMA_0_100) %>%
  write.csv("../data/outcome/spma_index.csv", row.names = FALSE)
```
