---
title: "cl_01_fame"
author: "Pragati Dahal, Yuanyuan Wen, Xiaoyi Zhao"
date: "2026-07-07"
output: html_document
---

## Purpose
The goal of this code here is to do separate the variables needed from the index

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load the data
```{r}
fame <- read_csv("../data/source/FAME_MasterData_Feb2026.csv")
```

# variables we need for the SPMA index + moderators
```{r}
spma_vars <- c(
  "all_farms",
  "all_sales",
  "annual_avg_estabs_count_naics_4244_grocery_and_related_product_wholesalers",
  "annual_avg_estabs_count_naics_4245_farm_product_raw_material_merch._whls.",
  "annual_avg_estabs_count_naics_44511_supermarkets_and_other_grocery_stores",
  "annual_avg_estabs_count_naics_4451_grocery_stores",
  "d2c_farms_pct",
  "d2c_sales_pct",
  "intermediated_farms_pct",
  "intermediated_sales_pct",
  "local_sales_pct",
  "number_colleges_universities",
  "number_csa",
  "number_farmersmarket",
  "number_foodhub",
  "number_healthcare_facilities",
  "number_k12_schools",
  "number_onfarmmarket",
  "overall_food_insecurity_rate",
  "pct_households_with_no_vehicle",
  "pct_laccess_lowi",
  "total_acres_operated",
  "valueadded_farms_pct"
)
```

quick check none of these got typo'd before we go further
```{r}
setdiff(spma_vars, unique(fame$variable_name))
```

```{r}
id_cols <- c("fips", "ansicode", "county_name", "state_name", "state_abbrev")
```

these variables come from different years in FAME (2019-2025, not one clean
panel year), so instead of hardcoding a year we just grab the latest year
available for each variable individually
```{r}
fame_latest <- fame %>%
  filter(variable_name %in% spma_vars) %>%
  group_by(fips, variable_name) %>%
  slice_max(year, n = 1) %>%
  ungroup()
```

keeping a lookup of which year each variable actually came from, want this
for the methods writeup later
```{r}
var_year_lookup <- fame_latest %>%
  distinct(variable_name, year) %>%
  arrange(variable_name)

var_year_lookup
```

```{r}
spma_index <- fame_latest %>%
  select(all_of(id_cols), variable_name, value) %>%
  pivot_wider(names_from = variable_name, values_from = value) %>%
  select(all_of(id_cols), all_of(spma_vars))
```

## Save output
```{r}
write_csv(spma_index, "../data/outcome/fame_variables.csv")
```
