---
title: "cl_01_farmersmarket_distance"
author: "Yuanyuan Wen"
date: "2026-07-07"
output: html_document
---
## Purpose

The goal of this code:
Identify the distance from the centroid of each county to the nearest farmers market in the farmers market directory, using county centroid data from hub_distance_full.csv (the output of cl_01_distance).


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Get the county centroids from cl_01_distance's output
```{r}
names_hub <- c("rowidx", "GEOID", "state", "county_name", "distance_nearest_hub_mile",
               "nearest_hub_name", "nearest_hub_id", "hub_county_geoid", "hub_county_name",
               "hub_products", "hub_desert", "centroid_x", "centroid_y")

county_centroid <- read_csv("../data/outcome/hub_distance_full.csv",
                            skip = 1, col_names = names_hub,
                            col_types = cols(.default = col_character())) %>%
  transmute(GEOID       = sprintf("%05d", as.integer(GEOID)),
            state, county_name,
            centroid_x  = as.numeric(gsub("[^0-9.eE+-]", "", centroid_x)),
            centroid_y  = as.numeric(gsub("[^0-9.eE+-]", "", centroid_y))) %>%
  # drop Alaska & Hawaii: they sit far outside EPSG:5070 (CONUS Albers), so distances
  # computed for them in 5070 are distorted (e.g. an inflated ~1000-mile Aleutians value).
  filter(!state %in% c("Alaska", "Hawaii")) %>%
  st_as_sf(coords = c("centroid_x", "centroid_y"), crs = 5070)   # 5070 = NAD83 Albers, in metres
```


### Load farmer's market data
Load the dataset from the farmers market directory and calculate the distance from the county centroid to each of the farmers market

note: if a farmers market does not exist in this county, we will map the distance to the closest farmers market around

```{r}
market_raw <- read_excel("../data/source/farmersmarket_2026-623143939.xlsx",
                         col_types = "text")

market_directory <- market_raw %>%
  select(listing_id, listing_name, location_x, location_y) %>%
  rename(lon = location_x, lat = location_y) %>%
  mutate(lon = as.numeric(lon),
         lat = as.numeric(lat)) %>%
  filter(!is.na(lon), !is.na(lat))

#glimpse(market_directory)
```


make the markets spatial, and look up which county each market sits in
```{r}
markets_sf <- st_as_sf(market_directory, coords = c("lon", "lat"), crs = 4326)

options(tigris_use_cache = TRUE)   # cache the download so it is not re-fetched every run
us_counties   <- counties(cb = TRUE, year = 2023, progress_bar = FALSE)
county_lookup <- us_counties %>%
  select(STATEFP, GEOID, county_name = NAME) %>%
  st_transform(4326)
markets_geo   <- st_join(markets_sf, county_lookup, join = st_within)
```

convert the xy coordinate of the markets into meters by 5070 so that we can get the distance from the centroid to the market
```{r}
market_proj <- st_transform(markets_sf, 5070)

#check whether two are in the same scale
st_crs(market_proj) == st_crs(county_centroid)
```
**identify the distance between the centroid and the market
```{r}
distance_market <- st_distance(county_centroid, market_proj)

# (number of counties) x (number of farmers markets)
dim(distance_market)
```
for each of the row in the distance matrix, identify the market that is closest to each of the county and the distance
```{r}
nearest_dist <- apply(distance_market, 1, min)
nearest_idx  <- apply(distance_market, 1, which.min)
```

in the county dataset, for each of the county centroid, identify the name and the distance to the closest farmers market
```{r}

farmersmarket_distance <- county_centroid %>%
  select(GEOID, state, county_name) %>%
  mutate(
    `distance_nearest_farmersmarket(mile)` = set_units(set_units(nearest_dist, "m"), "mi") %>% as.numeric(),
    nearest_farmersmarket_name = market_directory$listing_name[nearest_idx],
    nearest_farmersmarket_id   = market_directory$listing_id[nearest_idx])
```

Merge this with other farmers market information (which county the nearest market sits in)
```{r}
market_attrs <- markets_geo %>%
  st_drop_geometry() %>%
  select(listing_id,
         market_county_geoid = GEOID,
         market_county_name  = county_name)

farmersmarket_distance_full <- farmersmarket_distance %>%
  left_join(market_attrs, by = c("nearest_farmersmarket_id" = "listing_id")) %>%
  mutate(farmersmarket_desert = (`distance_nearest_farmersmarket(mile)` >= 20))

```

 look at the distribution of the farmers market
```{r}
farmersmarket_distance_full %>%
  st_drop_geometry() %>%
  group_by(state) %>%
  summarise(
    n_counties = n(),
    mean_mi    = mean(`distance_nearest_farmersmarket(mile)`),
    median_mi  = median(`distance_nearest_farmersmarket(mile)`),
    min_mi     = min(`distance_nearest_farmersmarket(mile)`),
    max_mi     = max(`distance_nearest_farmersmarket(mile)`)
  ) %>%
  arrange(desc(mean_mi))

```

### save the dataset

```{r}
write.csv(farmersmarket_distance_full, "../data/outcome/farmersmarket_distance_full.csv")
```
