---
title: "fig_02"
author: "Xiaoyi Zhao"
date: "2026-07-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose 
1. the purpose of this file is to geneat the county comparison plot 
2. the county comparison plot will compare the differnce in similar spma but different 
 composition 


load the library 
```{r}
library(tigris)    # downloads Census TIGER shapefiles (county boundaries)
library(sf)        # the workhorse: stores and manipulates spatial data
library(dplyr)     # regular data wrangling once things look like data frames

library(ggplot2)
library(mapview)
library(readxl)
library(readr)
library(units)

library(tidycensus)
#to make the spider plot
library(scales)

#devtools::install_github("ricardo-bion/ggradar")
library(ggradar)

```

step 1 : load the dataset 
```{r}
# add the statename, we load the statename from the tidycensus first
state_lookup <- fips_codes %>% 
  distinct(state_code, state_name)


merged_final <- read_csv("../data/outcome/merged_final_xz.csv") %>% 
  select(county_name_hub, county_fips, spma, direct_to_consumer, foodhub_inter, distance_nearest_hub_mile, all_sales, distance_nearest_farmersmarket_mile, all_sales, d2c_sales_pct, foodhub_share, share) %>% 
  rename(fruit_nut_share = share) %>% 
  arrange(fruit_nut_share) %>% 
  filter(county_fips != "22087",
         !is.na(fruit_nut_share)) %>% 
  mutate(state_code = substr(county_fips,1, 2)) %>% 
  left_join(state_lookup, by = "state_code") %>% 
  mutate(county_name_hub = paste0(county_name_hub,", ", state_name))


```
After, identifying, Brazos, Texas(48041), and Kandiyohi, Michigan (27067) , Michigan, both have relatively high spma, but Kandiyohi's direct to consumer index is much lower than that of Dane, but the advantage of Willamson is that the distance to the nearest foodhub is much smaller than that of Dane 

###Step 2, make the visualization 


**Okanogan, Washington v.s. Yakima **

```{r}

fig_2_Okanogan_Yakima <- merged_final %>%
  mutate(
    direct_to_consumer_share        = percent_rank(d2c_sales_pct),
    d2c = percent_rank(-distance_nearest_farmersmarket_mile),
    foodhub_inter             = percent_rank(foodhub_inter),
    hub_proximity             = percent_rank(-distance_nearest_hub_mile),
    foodhub_share = percent_rank(foodhub_share),
    spma = percent_rank(spma),
    fruit_nut_share = percent_rank(fruit_nut_share)
  ) %>%
  filter(county_fips %in%  c("53047", "53077")) %>%
  select(group = county_name_hub,
         `Direct To Consumer Share` = direct_to_consumer_share,
         `Close to Foodhub`= hub_proximity,
         `Close to Farmer's Market` = d2c,
         `Foodhub Intermediated Share` = foodhub_share,
         `SPMA` = spma,
         `Fruit Nut Share` = fruit_nut_share) %>%
  ggradar(font.radar = "sans",
          values.radar   = c("0", "50", "100"),
          group.colours  = c("#861F41", "#E5751F"),
          grid.label.size = 5,
          axis.label.size = 4)+
  labs(title = "How Foodhub Support Counties?") +
  theme_void() +
  theme(
    plot.title      = element_text(face = "bold", size = 14, hjust = 0),
    plot.subtitle   = element_text(size = 11, color = "grey30",
                                   margin = margin(b = 10)),
    plot.caption    = element_text(size = 8, color = "grey40", hjust = 0),
    legend.position = "bottom",
    legend.title    = element_text(size = 9),
    legend.text     = element_text(size = 8),
    plot.margin     = margin(20, 120, 20, 120)
  ) + coord_equal(clip = "off")

ggsave("../figures/fig_2_Okanogan_Yakima.pdf", plot = fig_2_Okanogan_Yakima, width = 12, height = 5)

```


**Ventura, CA v.s. Yakima , Washington**
```{r, fig.width = 12, fig.height = 5}
merged_final %>%
  filter(county_fips %in%  c("06111", "53077"))
```

```{r}

fig_2_Ventura_Yakima <- merged_final %>%
  mutate(
    direct_to_consumer_share        = percent_rank(d2c_sales_pct),
    d2c = percent_rank(-distance_nearest_farmersmarket_mile),
    foodhub_inter             = percent_rank(foodhub_inter),
    hub_proximity             = percent_rank(-distance_nearest_hub_mile),
    foodhub_share = percent_rank(foodhub_share),
    spma = percent_rank(spma),
    fruit_nut_share = percent_rank(fruit_nut_share)
  ) %>%
  filter(county_fips %in%  c("06111", "53077")) %>%
  select(group = county_name_hub,
         `Direct To Consumer Share` = direct_to_consumer_share,
         `Close to Foodhub`= hub_proximity,
         `Close to Farmer's Market` = d2c,
         `Foodhub Intermediated Share` = foodhub_share,
         `SPMA` = spma,
         `Fruit Nut Share` = fruit_nut_share) %>%
  mutate(group = recode(group,
                        "Brazos"  = "Brazos County, Texas",
                        "Kandiyohi" = "Kandiyohi County, Michigan")) %>%
  ggradar(font.radar = "sans",
          values.radar   = c("0", "50", "100"),
          group.colours  = c("#861F41", "#E5751F"),
          grid.label.size = 5,
          axis.label.size = 4)+
  labs(title = "How Foodhub Support Counties?") +
  theme_void() +
  theme(
    plot.title      = element_text(face = "bold", size = 14, hjust = 0),
    plot.subtitle   = element_text(size = 11, color = "grey30",
                                   margin = margin(b = 10)),
    plot.caption    = element_text(size = 8, color = "grey40", hjust = 0),
    legend.position = "bottom",
    legend.title    = element_text(size = 9),
    legend.text     = element_text(size = 8),
    plot.margin     = margin(20, 120, 20, 120)
  ) + coord_equal(clip = "off")

ggsave("../figures/fig_2_Ventura_Yakima.pdf", plot = fig_2_Ventura_Yakima, width = 12, height = 5)

```


**Monterey, CA v.s. Kern ,CA**

```{r}

fig_2_Monterey_Kern <- merged_final %>%
  mutate(
    direct_to_consumer_share        = percent_rank(d2c_sales_pct),
    d2c = percent_rank(-distance_nearest_farmersmarket_mile),
    foodhub_inter             = percent_rank(foodhub_inter),
    hub_proximity             = percent_rank(-distance_nearest_hub_mile),
    foodhub_share = percent_rank(foodhub_share),
    spma = percent_rank(spma),
    fruit_nut_share = percent_rank(fruit_nut_share)
  ) %>%
  filter(county_fips %in%  c("06053", "06029")) %>%
  select(group = county_name_hub,
         `Direct To Consumer Share` = direct_to_consumer_share,
         `Close to Foodhub`= hub_proximity,
         `Close to Farmer's Market` = d2c,
         `Foodhub Intermediated Share` = foodhub_share,
         `SPMA` = spma,
         `Fruit Nut Share` = fruit_nut_share) %>%
  ggradar(font.radar = "sans",
          values.radar   = c("0", "50", "100"),
          group.colours  = c("#861F41", "#E5751F"),
          grid.label.size = 5,
          axis.label.size = 4)+
  labs(title = "How Foodhub Support Counties?") +
  theme_void() +
  theme(
    plot.title      = element_text(face = "bold", size = 14, hjust = 0),
    plot.subtitle   = element_text(size = 11, color = "grey30",
                                   margin = margin(b = 10)),
    plot.caption    = element_text(size = 8, color = "grey40", hjust = 0),
    legend.position = "bottom",
    legend.title    = element_text(size = 9),
    legend.text     = element_text(size = 8),
    plot.margin     = margin(20, 120, 20, 120)
  ) + coord_equal(clip = "off")

fig_2_Monterey_Kern

ggsave("../figures/fig_2_Monterey_Kern.pdf", plot = fig_2_Monterey_Kern, width = 12, height = 5)
```


**Washington, Akansas v.s. Benton, Arkansas **

```{r}

fig_2_Washington_Benton <- merged_final %>%
  mutate(
    direct_to_consumer_share        = percent_rank(d2c_sales_pct),
    d2c = percent_rank(-distance_nearest_farmersmarket_mile),
    foodhub_inter             = percent_rank(foodhub_inter),
    hub_proximity             = percent_rank(-distance_nearest_hub_mile),
    foodhub_share = percent_rank(foodhub_share),
    spma = percent_rank(spma),
    fruit_nut_share = percent_rank(fruit_nut_share)
  ) %>%
  filter(county_fips %in%  c("05007", "05145")) %>%
  select(group = county_name_hub,
         `Direct To Consumer Share` = direct_to_consumer_share,
         `Close to Foodhub`= hub_proximity,
         `Close to Farmer's Market` = d2c,
         `Foodhub Intermediated Share` = foodhub_share,
         `SPMA` = spma,
         `Fruit Nut Share` = fruit_nut_share) %>%
  mutate(group = recode(group,
                        "Guadalupe"  = "Guadalupe County, TX",
                        "Guthrie" = "Guthrie County, TN")) %>%
  ggradar(font.radar = "sans",
          values.radar   = c("0", "50", "100"),
          group.colours  = c("#861F41", "#E5751F"),
          grid.label.size = 5,
          axis.label.size = 4)+
  labs(title = "How Foodhub Support Counties?") +
  theme_void() +
  theme(
    plot.title      = element_text(face = "bold", size = 14, hjust = 0),
    plot.subtitle   = element_text(size = 11, color = "grey30",
                                   margin = margin(b = 10)),
    plot.caption    = element_text(size = 8, color = "grey40", hjust = 0),
    legend.position = "bottom",
    legend.title    = element_text(size = 9),
    legend.text     = element_text(size = 8),
    plot.margin     = margin(20, 120, 20, 120)
  ) + coord_equal(clip = "off")

ggsave("../figures/fig_2_Washington_Benton.pdf", plot = fig_2_Washington_Benton, width = 12, height = 5)

```
**Guadalupe, TX vs Cass, MN( Minnisota)**

```{r}

fig_2_Guadalupe_Cass <- merged_final %>%
  mutate(
    direct_to_consumer_share        = percent_rank(d2c_sales_pct),
    d2c = percent_rank(-distance_nearest_farmersmarket_mile),
    foodhub_inter             = percent_rank(foodhub_inter),
    hub_proximity             = percent_rank(-distance_nearest_hub_mile),
    foodhub_share = percent_rank(foodhub_share),
    spma = percent_rank(spma),
    fruit_nut_share = percent_rank(fruit_nut_share)
  ) %>%
  filter(county_fips %in%  c("48187", "26027")) %>%
  select(group = county_name_hub,
         `Direct To Consumer Share` = direct_to_consumer_share,
         `Close to Foodhub`= hub_proximity,
         `Close to Farmer's Market` = d2c,
         `Foodhub Intermediated Share` = foodhub_share,
         `SPMA` = spma,
         `Fruit Nut Share` = fruit_nut_share) %>%
  mutate(group = recode(group,
                        "Guadalupe"  = "Guadalupe County, TX",
                        "Cass" = "Cass County, MN")) %>%
  ggradar(font.radar = "sans",
          values.radar   = c("0", "50", "100"),
          group.colours  = c("#861F41", "#E5751F"),
          grid.label.size = 5,
          axis.label.size = 4)+
  labs(title = "How Foodhub Support Counties?") +
  theme_void() +
  theme(
    plot.title      = element_text(face = "bold", size = 14, hjust = 0),
    plot.subtitle   = element_text(size = 11, color = "grey30",
                                   margin = margin(b = 10)),
    plot.caption    = element_text(size = 8, color = "grey40", hjust = 0),
    legend.position = "bottom",
    legend.title    = element_text(size = 9),
    legend.text     = element_text(size = 8),
    plot.margin     = margin(20, 120, 20, 120)
  ) + coord_equal(clip = "off")

ggsave("../figures/fig_2_Guadalupe_Cass.pdf", plot = fig_2_Guadalupe_Cass, width = 12, height = 5)

```


**Grawford, GA vs Camden, GA**
```{r}

fig_2_Grawford_Camden <- merged_final %>%
  mutate(
    direct_to_consumer_share        = percent_rank(d2c_sales_pct),
    d2c = percent_rank(-distance_nearest_farmersmarket_mile),
    foodhub_inter             = percent_rank(foodhub_inter),
    hub_proximity             = percent_rank(-distance_nearest_hub_mile),
    foodhub_share = percent_rank(foodhub_share),
    spma = percent_rank(spma),
    fruit_nut_share = percent_rank(fruit_nut_share)
  ) %>%
  filter(county_fips %in%  c("13079", "13039")) %>%
  select(group = county_name_hub,
         `Direct To Consumer Share` = direct_to_consumer_share,
         `Close to Foodhub`= hub_proximity,
         `Close to Farmer's Market` = d2c,
         `Foodhub Intermediated Share` = foodhub_share,
         `SPMA` = spma,
         `Fruit Nut Share` = fruit_nut_share) %>%
  ggradar(font.radar = "sans",
          values.radar   = c("0", "50", "100"),
          group.colours  = c("#861F41", "#E5751F"),
          grid.label.size = 5,
          axis.label.size = 4)+
  labs(title = "How Foodhub Support Counties?") +
  theme_void() +
  theme(
    plot.title      = element_text(face = "bold", size = 14, hjust = 0),
    plot.subtitle   = element_text(size = 11, color = "grey30",
                                   margin = margin(b = 10)),
    plot.caption    = element_text(size = 8, color = "grey40", hjust = 0),
    legend.position = "bottom",
    legend.title    = element_text(size = 9),
    legend.text     = element_text(size = 8),
    plot.margin     = margin(20, 120, 20, 120)
  ) + coord_equal(clip = "off")


```
```{r}
ggsave("../figures/fig_2_Grawford_Camden.pdf", plot = fig_2_Grawford_Camden, width = 12, height = 5)

```

**Franklin, WA vs. Chilton, Alabama**

```{r}

fig_2_Franklin_Chilton <- merged_final %>%
  mutate(
    direct_to_consumer_share        = percent_rank(d2c_sales_pct),
    d2c = percent_rank(-distance_nearest_farmersmarket_mile),
    foodhub_inter             = percent_rank(foodhub_inter),
    hub_proximity             = percent_rank(-distance_nearest_hub_mile),
    foodhub_share = percent_rank(foodhub_share),
    spma = percent_rank(spma),
    fruit_nut_share = percent_rank(fruit_nut_share)
  ) %>%
  filter(county_fips %in%  c("53021", "01021")) %>%
  select(group = county_name_hub,
         `Direct To Consumer Share` = direct_to_consumer_share,
         `Close to Foodhub`= hub_proximity,
         `Close to Farmer's Market` = d2c,
         `Foodhub Intermediated Share` = foodhub_share,
         `SPMA` = spma,
         `Fruit Nut Share` = fruit_nut_share) %>%
  ggradar(font.radar = "sans",
          values.radar   = c("0", "50", "100"),
          group.colours  = c("#861F41", "#E5751F"),
          grid.label.size = 5,
          axis.label.size = 4)+
  labs(title = "How Foodhub Support Counties?") +
  theme_void() +
  theme(
    plot.title      = element_text(face = "bold", size = 14, hjust = 0),
    plot.subtitle   = element_text(size = 11, color = "grey30",
                                   margin = margin(b = 10)),
    plot.caption    = element_text(size = 8, color = "grey40", hjust = 0),
    legend.position = "bottom",
    legend.title    = element_text(size = 9),
    legend.text     = element_text(size = 8),
    plot.margin     = margin(20, 120, 20, 120)
  ) + coord_equal(clip = "off")

ggsave("../figures/fig_2_Franklin_Chilton.pdf", plot = fig_2_Franklin_Chilton, width = 12, height = 5)

```

**Greenville County, SC  v.s. Buncombe County, NC**

```{r}

fig_2_Greenville_Buncombe <- merged_final %>%
  mutate(
    direct_to_consumer_share        = percent_rank(d2c_sales_pct),
    d2c = percent_rank(-distance_nearest_farmersmarket_mile),
    foodhub_inter             = percent_rank(foodhub_inter),
    hub_proximity             = percent_rank(-distance_nearest_hub_mile),
    foodhub_share = percent_rank(foodhub_share),
    spma = percent_rank(spma),
    fruit_nut_share = percent_rank(fruit_nut_share)
  ) %>%
  filter(county_fips %in%  c("45045", "37021")) %>%
  select(group = county_name_hub,
         `Direct To Consumer Share` = direct_to_consumer_share,
         `Close to Foodhub`= hub_proximity,
         `Close to Farmer's Market` = d2c,
         `Foodhub Intermediated Share` = foodhub_share,
         `SPMA` = spma,
         `Fruit Nut Share` = fruit_nut_share) %>%
  ggradar(font.radar = "sans",
          values.radar   = c("0", "50", "100"),
          group.colours  = c("#861F41", "#E5751F"),
          grid.label.size = 5,
          axis.label.size = 4)+
  labs(title = "How Foodhub Support Counties?") +
  theme_void() +
  theme(
    plot.title      = element_text(face = "bold", size = 14, hjust = 0),
    plot.subtitle   = element_text(size = 11, color = "grey30",
                                   margin = margin(b = 10)),
    plot.caption    = element_text(size = 8, color = "grey40", hjust = 0),
    legend.position = "bottom",
    legend.title    = element_text(size = 9),
    legend.text     = element_text(size = 8),
    plot.margin     = margin(20, 120, 20, 120)
  ) + coord_equal(clip = "off")

fig_2_Greenville_Buncombe
ggsave("../figures/fig_2_Greenville_Buncombe.pdf", plot = fig_2_Greenville_Buncombe, width = 12, height = 5)

```


**Marion County County, FL  v.s. Okeechobee County, FL**

```{r}

fig_2_Marion_Okeechobee <- merged_final %>%
  mutate(
    direct_to_consumer_share        = percent_rank(d2c_sales_pct),
    d2c = percent_rank(-distance_nearest_farmersmarket_mile),
    foodhub_inter             = percent_rank(foodhub_inter),
    hub_proximity             = percent_rank(-distance_nearest_hub_mile),
    foodhub_share = percent_rank(foodhub_share),
    spma = percent_rank(spma),
    fruit_nut_share = percent_rank(fruit_nut_share)
  ) %>%
  filter(county_fips %in%  c("12083", "12093")) %>%
  select(group = county_name_hub,
         `Direct To Consumer Share` = direct_to_consumer_share,
         `Close to Foodhub`= hub_proximity,
         `Close to Farmer's Market` = d2c,
         `Foodhub Intermediated Share` = foodhub_share,
         `SPMA` = spma,
         `Fruit Nut Share` = fruit_nut_share) %>%
  ggradar(font.radar = "sans",
          values.radar   = c("0", "50", "100"),
          group.colours  = c("#861F41", "#E5751F"),
          grid.label.size = 5,
          axis.label.size = 4)+
  labs(title = "How Foodhub Support Counties?") +
  theme_void() +
  theme(
    plot.title      = element_text(face = "bold", size = 14, hjust = 0),
    plot.subtitle   = element_text(size = 11, color = "grey30",
                                   margin = margin(b = 10)),
    plot.caption    = element_text(size = 8, color = "grey40", hjust = 0),
    legend.position = "bottom",
    legend.title    = element_text(size = 9),
    legend.text     = element_text(size = 8),
    plot.margin     = margin(20, 120, 20, 120)
  ) + coord_equal(clip = "off")

fig_2_Marion_Okeechobee
ggsave("../figures/fig_2_Marion_Okeechobee.pdf", plot = fig_2_Marion_Okeechobee, width = 12, height = 5)

```