---
title: "fig_01_main"
author: "Pragati Dahal, Xiaoyi Zhao"
date: "2026-07-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose

The goal of this file to build the first candidate of the visualation, 

this will be by state level, and we will plot the market accessibility by state 


load the libraries
```{r}
library(tigris)
library(sf)
library(dplyr)
library(ggplot2)
library(readxl)
library(readr)
library(ggnewscale)
options(tigris_use_cache = TRUE)
```


load the dataset we merged 
```{r}
merged_final <- read_csv("../data/outcome/merged_final_xz.csv") %>% 
  select(state, spma) %>% 
  group_by(state) %>% 
  summarize(
    spma = mean(spma, na.rm = TRUE)
  ) %>% 
  mutate(state = tolower(state))
  # we ignore the missing data here 
```

Plot the figure : 

```{r}

# first, we load the us state level map with simple boundaries

state_map <- states(cb = TRUE) %>% 
  filter(!STUSPS %in% c("AK", "HI", "PR", "GU", "VI", "AS", "MP"))
```
step 2: join the index with the map 
```{r}
index_map <- state_map %>% 
  mutate(state_lower = tolower(NAME)) %>%  # to match with our statename
  left_join(merged_final, by = c("state_lower" = "state"))
```


step 3: graph the first figure:

```{r}
index_map <- index_map %>%
  mutate(spma_scaled = 100 * spma / max(spma, na.rm = TRUE))

ggplot2::ggplot(index_map) +
  geom_sf(aes(fill = spma_scaled), color = "grey70", linewidth = 0.2) +
   scale_fill_distiller(
  palette   = "Reds",
  direction = 1,                    # 1 = darker for larger
  trans     = "log10",
  labels    = scales::label_number(accuracy = 0.1),
  na.value  = "grey88",
  name      = "SPMA\n(log scale)"
) +
  coord_sf(crs = 5070) +           # Albers Equal Area projection
  labs(
    title    = "Small-Producer Market Access (SPMA) Index in State Means",
    subtitle = "Higher = better weighted access to farmers markets & food hubs",
    caption  = "Gray = no data. Index rescaled to 0–100 relative to the highest state. Pink dots = Location of Foodhubs."
  ) +
  theme_void() +
  theme(
    plot.title      = element_text(face = "bold", size = 14, hjust = 0),
    plot.subtitle   = element_text(size = 11, color = "grey30",
                                   margin = margin(b = 10)),
    plot.caption    = element_text(size = 8, color = "grey40", hjust = 0),
    legend.position = c(0.92, 0.30),
    legend.title    = element_text(size = 9),
    legend.text     = element_text(size = 8),
    plot.margin     = margin(10, 10, 10, 10)
  )
```
step 4: overlay this with the hot splot of foodhub 

load the foodhub locations
```{r}
foodhub_raw <- read_excel("../data/source/foodhub_2026-75161420.xlsx") 

foodhub_directory <- foodhub_raw %>% 
  select(listing_id, listing_name, location_x, location_y, products) %>% 
  rename(lon = location_x, lat = location_y) %>% 
  mutate(lon = as.numeric(lon), 
         lat = as.numeric(lat)) %>% 
  filter(!is.na(lon), !is.na(lat)) 

hubs_sf <- st_as_sf(foodhub_directory, coords = c("lon", "lat"), crs = 4326)

us_counties <- counties(cb = TRUE, year = 2023)

county_lookup <- us_counties %>%
  select(STATEFP, GEOID, county_name = NAME) %>%   # select() can rename in-place
  st_transform(4326) 

hubs_geo <- st_join(hubs_sf, county_lookup, join = st_within)

hubs_pts <- st_transform(hubs_geo) %>% 
  filter(!STATEFP %in% c("02", "15"))                           # exclude Hawaii and Alaska

```
# --- crop groups: state-level, sourced from USDA ERS FTS-384 (March 2026) ----
```{r}
crop_lu <- tibble::tribble(
  ~state_name,      ~crop_group,
  "WASHINGTON",     "Apples & Pears",      # 2nd in US fruit/nut cash receipts, p.2
  "OREGON",         "Apples & Pears",
  "CALIFORNIA",     "Tree Nuts & Grapes",  # ~all US almonds & pistachios, p.27
  "WISCONSIN",      "Berries",
  "MASSACHUSETTS",  "Berries",
  "MICHIGAN",       "Berries",
  "MAINE",          "Berries",             # wild lowbush blueberry, p.14
  "FLORIDA",        "Peaches & Citrus",    # strawberries 56% of FL receipts, p.13
  "GEORGIA",        "Peaches & Citrus",    # leading watermelon producer, p.23
  "SOUTH CAROLINA", "Peaches & Citrus",
  "TEXAS",          "Peaches & Citrus"     # grapefruit, p.9
)

fruit <- read_csv("../data/outcome/fruitnut_share.csv") %>%
  transmute(county_fips = sprintf("%05d", as.integer(fips)),
            state_name  = `state_name.x`,
            fruitnut_sale) %>%
  filter(!is.na(fruitnut_sale)) %>%
  inner_join(crop_lu, by = "state_name")

crop_pts <- counties(cb = TRUE, year = 2023, progress_bar = FALSE) %>%
  filter(!STATEFP %in% c("02","15","60","66","69","72","78")) %>%
  transmute(county_fips = GEOID) %>%
  inner_join(fruit, by = "county_fips") %>%
  group_by(crop_group) %>%
  slice_max(fruitnut_sale, n = 15) %>%
  ungroup() %>%
  st_transform(5070) %>%
  st_point_on_surface()

pal <- c("Apples & Pears"     = "#2E7D32",   # deep green
         "Tree Nuts & Grapes" = "#6A1B9A",   # purple
         "Berries"            = "#C2185B",   # magenta
         "Peaches & Citrus"   = "#F57F17")   # amber
```         

```{r, fig.width= 12, fig.height = 5}
fig_1 <- ggplot(index_map) +
  geom_sf(aes(fill = spma_scaled), color = "grey75", linewidth = 0.2) +
  scale_fill_distiller(
    palette   = "Blues",
    direction = 1,
    trans     = "log10",
    labels    = scales::label_number(accuracy = 0.01),
    na.value  = "grey90",
    name      = "SPMA\n(log scale)",
    guide     = guide_colorbar(order = 1)
  ) +
  new_scale_fill() +
  geom_sf(data = hubs_pts, shape = 21, fill = "grey85",
          color = "grey50", size = 2.0, stroke = 0.25, alpha = 0.5) +
  geom_sf(data = crop_pts, aes(fill = crop_group, size = fruitnut_sale),
          shape = 21, color = "white", stroke = 0.4, alpha = 0.95) +
  scale_fill_manual(values = pal, name = "Crop group",
                    guide = guide_legend(order = 2,
                                         override.aes = list(size = 4))) +
  scale_size_continuous(range = c(1, 5), trans = "sqrt", guide = "none") +
  coord_sf(crs = 5070) +
  labs(
    title    = "Small-Producer Market Access (SPMA) Index in State Medians",
    subtitle = "Higher = better weighted access to farmers markets & food hubs",
    caption  = paste0(
      "Pale dots = food hub locations.\n",
      "Colored dots = 15 highest fruit/nut sales counties per crop group ",
      "(2022 Census of Agriculture), sized by sales value.\n",
      "Crop groups assigned at state level from USDA ERS, Fruit and Tree Nuts ",
      "Outlook (FTS-384, March 2026)."
    )
  ) +
  theme_void() +
  theme(
    plot.title      = element_text(face = "bold", size = 14, hjust = 0),
    plot.subtitle   = element_text(size = 11, color = "grey30",
                                   margin = margin(b = 10)),
    plot.caption    = element_text(size = 7.5, color = "grey40", hjust = 0),
    legend.position = "right",
    legend.title    = element_text(size = 9),
    legend.text     = element_text(size = 8),
    plot.margin     = margin(10, 10, 10, 10)
  )
```

```{r}
ggsave("../figures/fig_01_xz_fruits.pdf", plot = fig_1, width = 12, height = 5)
fig_1
```

