---
title: "Figure 02: County Level Comparison"
author: "Xiaoyi Zhao, Pragati Dahal, Yuanyuan Wen"
date: "2026-07-12"
output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  warning = FALSE,   # hides warnings
  message = FALSE,   # hides package startup messages, etc.
  echo = FALSE
)
```

## Purpose 
1. the purpose of this file is to generate the county comparison plot 
2. The counties are choose are Greenville, SC vs Lee, GA 
3. We will plot county comparison first
4. Then, we will plot the whatif case


### Step 1: load the library and the merged dataset 
\begin{itemize}
  \item The index and the factors represented in the spider graph are percentile
  \item Since local intermediate sale percentage has a lot of missing values, this county comparison use $(1-(\text{direct to consumer sales' share}))$ as the the estimation of the food hub intermediated sale 
\end{itemize}
```{r include=FALSE}
library(tigris)    # downloads Census TIGER shapefiles (county boundaries)
library(sf)        # the workhorse: stores and manipulates spatial data
library(dplyr)     # regular data wrangling once things look like data frames

library(ggplot2)
library(mapview)
library(readxl)
library(readr)
library(units)

library(tidycensus)
#to make the spider plot
library(scales)

#devtools::install_github("ricardo-bion/ggradar")
library(ggradar)

library(showtext)
library(sysfonts)
library(patchwork)
library(ggtext)

```

Specify the font that we want to use for the visualization
```{r}
# set DM Serif Display as the font of title 
font_add_google(
  name = "DM Serif Display",
  family = "title_font"
)

# set DM Sans as the font of the subtitle and body text if we add notes
font_add_google(
  name = "DM Sans",
  family = "body_font"
)

# we need to keep this open so that the figure can show the 
# customized fonts above
showtext::showtext_auto(TRUE)

```

```{r include=FALSE}
# add the statename, we load the statename from the tidycensus first
state_lookup <- fips_codes %>% 
  distinct(state_code, state_name)
  
# load the merged dataset and respecify the names 
merged_final <- read_csv("../data/outcome/merged_final_xz.csv") %>% 
  select(county_name_hub, county_fips, spma, direct_to_consumer, foodhub_inter, distance_nearest_hub_mile, all_sales, distance_nearest_farmersmarket_mile, all_sales, d2c_sales_pct, foodhub_share, share) %>% 
  rename(fruit_nut_share = share) %>% 
  arrange(fruit_nut_share) %>% 
  filter(county_fips != "22087",
         !is.na(fruit_nut_share)) %>% 
  mutate(state_code = substr(county_fips,1, 2)) %>% 
  left_join(state_lookup, by = "state_code") %>% 
  mutate(county_name_hub = paste0(county_name_hub,", ", state_name))


```
\newpage

### Step 2:  County Comparison

**Greenville, SC  v.s. Lee County,GA**

The Fips we identified are 
```{r echo=FALSE, fig.width= 12, fig.height= 5}


fig_2_Greenville_Lee <- merged_final %>%
  mutate(
    direct_to_consumer_share        = percent_rank(d2c_sales_pct),
    d2c = percent_rank(-distance_nearest_farmersmarket_mile),
    foodhub_inter             = percent_rank(foodhub_inter),
    hub_proximity             = percent_rank(-distance_nearest_hub_mile),
    foodhub_share = percent_rank(foodhub_share),
    spma = percent_rank(spma),
    fruit_nut_share = percent_rank(fruit_nut_share)
  ) %>%
  filter(county_fips %in%  c("45045", "13177")) %>%
  select(group = county_name_hub,
         `Direct To \nConsumer Share` = direct_to_consumer_share,
         `Close to \nFOODHUB`= hub_proximity,
         `Close to \nFarmer's Market` = d2c,
         `Foodhub Intermediated Share` = foodhub_share,
         `SPMA` = spma,
         `Fruit Nut Share` = fruit_nut_share) %>%
  ggradar(font.radar = "roboto",
          values.radar   = c("0", "50", "100"),
          group.colours  = c("#3182BD", "#FFD23F"),
          grid.label.size = 5,
          axis.label.size = 4,
          axis.label.offset = 1.2,
          gridline.min.colour = "#BFDBFE",
          gridline.mid.colour = "#60A5FA",
          gridline.max.colour = "#1E3A8A") +
  theme_void(base_family = "body_font") +
  labs(
    title = "From Orchard to Opportunity",
    subtitle = "Food Hubs and Small Farm Market Access at the County Level") +
  theme(
    plot.background = element_rect(
      fill = "#FAF8F2",
      color = NA
    ),
    panel.background = element_rect(
      fill = "#FAF8F2",
      color = NA
    ),

    plot.title = element_text(
      family = "title_font",
      size = 25,
      lineheight = 0.88,
      color = "#123B65",
      face = "plain",
      hjust = 0,
      margin = margin(b = 10)
    ),

    plot.subtitle = element_text(
      family = "body_font",
      size = 10,
      lineheight = 1.2,
      color = "#4B5D6D",
      hjust = 0,
      margin = margin(b = 12)
    ),

    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.width = unit(2.5, "cm"),
    legend.key.height = unit(0.28, "cm"),

    plot.margin = margin(
  t = 10,
  r = 10,
  b = 10,
  l = 10
)
) + coord_equal(clip = "off")

fig_2_Greenville_Lee

```

### Step 3: Plot the graph over its whatif case


