---
title: "Figure 02: County Level Comparison"
author: "Xiaoyi Zhao, Pragati Dahal, Yuanyuan Wen"
date: "2026-07-12"
output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  warning = FALSE,   # hides warnings
  message = FALSE,   # hides package startup messages, etc.
  echo = FALSE
)
```

## Purpose 
1. the purpose of this file is to generate the county comparison plot 
2. The counties are choose are Greenville, SC vs Lee, GA 
3. We will plot county comparison first
4. Then, we will plot the whatif case

\begin{itemize}
  \item The index and the factors represented in the spider graph are percentile
  \item Since local intermediate sale percentage has a lot of missing values, this county comparison use $(1-(\text{direct to consumer sales' share}))$ as the the estimation of the food hub intermediated sale 
\end{itemize}

### Step 1: specify the font

```{r}
# set DM Serif Display as the font of title 
font_add_google(
  name = "DM Serif Display",
  family = "title_font"
)

# set DM Sans as the font of the subtitle and body text if we add notes
font_add_google(
  name = "DM Sans",
  family = "body_font"
)


```

```{r include=FALSE}
# add the statename, we load the statename from the tidycensus first
state_lookup <- fips_codes %>% 
  distinct(state_code, state_name)
  
# load the merged dataset and respecify the names 
merged_final <- read_csv("../data/outcome/spma_index.csv") %>% 
  select(county_name, county_fips, SPMA_0_100, distance_nearest_hub_mile, distance_nearest_farmersmarket_mile, d2c_sales_pct, intermediated_sales_pct_derived, share) %>%
  rename(spma = SPMA_0_100,
    fruit_nut_share = share,
    foodhub_share = intermediated_sales_pct_derived) %>% 
  arrange(fruit_nut_share) %>% 
  filter(county_fips != "22087",
         !is.na(fruit_nut_share)) %>% 
  mutate(state_code = substr(county_fips,1, 2)) %>% 
  left_join(state_lookup, by = "state_code") %>% 
  mutate(county_name_hub = paste0(county_name,", ", state_name))

```
\newpage

### Step 2:  County Comparison

**Greenville, SC  v.s. Lee County,GA**

The Fips we identified are 
```{r echo=FALSE, fig.width= 12, fig.height= 5}

# we need to keep this open so that the figure can show the 
# customized fonts above
showtext::showtext_auto(TRUE)

fig_2_Greenville_Lee <- merged_final %>%
  mutate(
    direct_to_consumer_share        = percent_rank(d2c_sales_pct),
    d2c = percent_rank(-distance_nearest_farmersmarket_mile),
    hub_proximity             = percent_rank(-distance_nearest_hub_mile),
    foodhub_share = percent_rank(foodhub_share),
    spma = percent_rank(spma),
    fruit_nut_share = percent_rank(fruit_nut_share)
  ) %>%
  filter(county_fips %in%  c("45045", "13177")) %>%
  select(group = county_name_hub,
         `Direct To Consumer Share` = direct_to_consumer_share,
         `Close to \nFOOD HUB`= hub_proximity,
         `Close to \nFarmer's \nMarket` = d2c,
         `Food hub Intermediated Share` = foodhub_share,
         `SPMA` = spma,
         `Fruit Nut Share` = fruit_nut_share) %>%
  ggradar(font.radar = "roboto",
          values.radar   = c("0", "50", "100"),
          group.colours  = c("#013e75", "#b05b24"),
          grid.label.size = 5,
          axis.label.size = 4,
          axis.label.offset = 1.2,
          gridline.min.colour = "#BFDBFE",
          gridline.mid.colour = "#60A5FA",
          gridline.max.colour = "#1E3A8A") +
  theme_void(base_family = "body_font") +
  theme(
    plot.background = element_rect(
      fill = "#FAF8F2",
      color = NA
    ),
    panel.background = element_rect(
      fill = "#FAF8F2",
      color = NA
    ),

    plot.title = element_text(
      family = "title_font",
      size = 25,
      lineheight = 0.88,
      color = "#123B65",
      face = "plain",
      hjust = 0,
      margin = margin(b = 10)
    ),

    plot.subtitle = element_text(
      family = "body_font",
      size = 10,
      lineheight = 1.2,
      color = "#4B5D6D",
      hjust = 0,
      margin = margin(b = 12)
    ),

    legend.position = "left",
    legend.direction = "vertical",
    legend.key.width = unit(2, "cm"),
    legend.key.height = unit(0.14, "cm"),

    plot.margin = margin(
  t = 10,
  r = 10,
  b = 10,
  l = 10
)
) + coord_equal(clip = "off")

fig_2_Greenville_Lee

```

**Yakima, WA  v.s. Okanogan, WA**

The Fips we identified are 
```{r echo=FALSE, fig.width= 12, fig.height= 5}


fig_2_Yakima_Okanogan <- merged_final %>%
  mutate(
    direct_to_consumer_share        = percent_rank(d2c_sales_pct),
    d2c = percent_rank(-distance_nearest_farmersmarket_mile),
    hub_proximity             = percent_rank(-distance_nearest_hub_mile),
    foodhub_share = percent_rank(foodhub_share),
    spma = percent_rank(spma),
    fruit_nut_share = percent_rank(fruit_nut_share)
  ) %>%
  filter(county_fips %in%  c("53077", "53047")) %>%
  select(group = county_name_hub,
         `Direct To Consumer Share` = direct_to_consumer_share,
         `Close to \nFOOD HUB`= hub_proximity,
         `Close to \nFarmer's \nMarket` = d2c,
         `Food hub Intermediated Share` = foodhub_share,
         `SPMA` = spma,
         `Fruit Nut Share` = fruit_nut_share) %>%
  ggradar(font.radar = "roboto",
          values.radar   = c("0", "50", "100"),
          group.colours  = c("#013e75", "#b05b24"),
          grid.label.size = 5,
          axis.label.size = 4,
          axis.label.offset = 1.2,
          gridline.min.colour = "#BFDBFE",
          gridline.mid.colour = "#60A5FA",
          gridline.max.colour = "#1E3A8A") +
  theme_void(base_family = "body_font")  +
  theme(
    plot.background = element_rect(
      fill = "#FAF8F2",
      color = NA
    ),
    panel.background = element_rect(
      fill = "#FAF8F2",
      color = NA
    ),

    plot.title = element_text(
      family = "title_font",
      size = 25,
      lineheight = 0.88,
      color = "#123B65",
      face = "plain",
      hjust = 0,
      margin = margin(b = 10)
    ),

    plot.subtitle = element_text(
      family = "body_font",
      size = 10,
      lineheight = 1.2,
      color = "#4B5D6D",
      hjust = 0,
      margin = margin(b = 12)
    ),

    legend.position = "left",
    legend.direction = "vertical",
    legend.key.width = unit(2, "cm"),
    legend.key.height = unit(0.14, "cm"),

    plot.margin = margin(
  t = 10,
  r = 10,
  b = 10,
  l = 10
)
) + coord_equal(clip = "off")

fig_2_Yakima_Okanogan

```

### Step 3: Plot the graph over its whatif case 1

load the simulated data first 
```{r}

merged_whatif_lee <- read_csv("../data/outcome/spma_index_greenville_lee.csv") %>% 
  mutate(distance_nearest_hub_mile_exchange = rev(distance_nearest_hub_mile)) %>% 
  filter(county_fips == 13177) %>% 
  select(county_name, county_fips, SPMA_exchange_0_100, distance_nearest_hub_mile_exchange, distance_nearest_farmersmarket_mile, d2c_sales_pct, share) %>%
  mutate(foodhub_share = if_else(d2c_sales_pct >= 1, NA, 100 - d2c_sales_pct)) %>% 
  rename(distance_nearest_hub_mile = distance_nearest_hub_mile_exchange,
         spma = SPMA_exchange_0_100,
         fruit_nut_share = share) %>% 
  mutate(state_code = substr(county_fips,1, 2)) %>% 
  left_join(state_lookup, by = "state_code") %>% 
  mutate(county_name_hub = paste0(county_name,", ", state_name, " (With Closer Foodhub)"))

merged_final_whatif_lee <- rbind(merged_final %>% filter(county_fips!= 13177), merged_whatif_lee)


```

```{r echo=FALSE, fig.width= 12, fig.height= 5}


fig_2_lee <- merged_final_whatif_lee %>%
  mutate(
    direct_to_consumer_share        = percent_rank(d2c_sales_pct),
    d2c = percent_rank(-distance_nearest_farmersmarket_mile),
    hub_proximity             = percent_rank(-distance_nearest_hub_mile),
    foodhub_share = percent_rank(foodhub_share),
    spma = percent_rank(spma),
    fruit_nut_share = percent_rank(fruit_nut_share)
  ) %>%
  filter(county_fips %in% c("13177", "45045")) %>%
  select(group = county_name_hub,
         `Direct To Consumer Share` = direct_to_consumer_share,
         `Close to \nFOOD HUB`= hub_proximity,
         `Close to \nFarmer's \nMarket` = d2c,
         `Food hub Intermediated Share` = foodhub_share,
         `SPMA` = spma,
         `Fruit Nut Share` = fruit_nut_share) %>%
  ggradar(font.radar = "roboto",
          values.radar   = c("0", "50", "100"),
          group.colours  = c("#013e75", "#a42423"),
          grid.label.size = 5,
          axis.label.size = 4,
          axis.label.offset = 1.2,
           gridline.min.colour = "#BFDBFE",
          gridline.mid.colour = "#60A5FA",
          gridline.max.colour = "#1E3A8A") +
  theme_void(base_family = "body_font") +
  theme(
    plot.background = element_rect(
      fill = "#FAF8F2",
      color = NA
    ),
    panel.background = element_rect(
      fill = "#FAF8F2",
      color = NA
    ),
    plot.title = element_text(
      family = "title_font",
      size = 20,
      lineheight = 0.88,
      color = "#15803D",
      face = "plain",
      hjust = 0,
      margin = margin(b = 8)
    ),

    plot.subtitle = element_text(
      family = "title_font",
      size = 20,
      lineheight = 1.2,
      color = "#15803D",
      hjust = 0,
      margin = margin(b = 12,
                      l = 50)
    ),

    legend.position = "right",
    legend.direction = "vertical",
    legend.key.width = unit(2, "cm"),
    legend.key.height = unit(0.14, "cm"),

    plot.margin = margin(
  t = 10,
  r = 10,
  b = 10,
  l = 10
)
) + coord_equal(clip = "off") 

fig_2_lee

```



### Step 4: Plot the graph over its whatif case 2: Yakima vs Okanogan in WA

load the simulated data first 
```{r}

merged_whatif_yakima <- read_csv("../data/outcome/spma_index_yakima_okanogan.csv") %>% 
  mutate(distance_nearest_hub_mile_exchange = rev(distance_nearest_hub_mile)) %>% 
  filter(county_fips == 53077) %>% 
  select(county_name, county_fips, SPMA_exchange_0_100, distance_nearest_hub_mile_exchange, distance_nearest_farmersmarket_mile, d2c_sales_pct, share) %>%
  mutate(foodhub_share = if_else(d2c_sales_pct >= 1, NA, 100 - d2c_sales_pct)) %>% 
  rename(distance_nearest_hub_mile = distance_nearest_hub_mile_exchange,
         spma = SPMA_exchange_0_100,
         fruit_nut_share = share) %>% 
  mutate(state_code = substr(county_fips,1, 2)) %>% 
  left_join(state_lookup, by = "state_code") %>% 
  mutate(county_name_hub = paste0(county_name,", ", state_name, " (With Closer Foodhub)"))

merged_final_whatif_yakima <- rbind(merged_final %>% filter(county_fips != 53077), merged_whatif_yakima)


```


```{r echo=FALSE, fig.width= 12, fig.height= 5}


fig_2_yakima <- merged_final_whatif_yakima %>%
  mutate(
    direct_to_consumer_share        = percent_rank(d2c_sales_pct),
    d2c = percent_rank(-distance_nearest_farmersmarket_mile),
    hub_proximity             = percent_rank(-distance_nearest_hub_mile),
    foodhub_share = percent_rank(foodhub_share),
    spma = percent_rank(spma),
    fruit_nut_share = percent_rank(fruit_nut_share)
  ) %>%
  filter(county_fips %in% c("53077", "53047")) %>%
  select(group = county_name_hub,
         `Direct To Consumer Share` = direct_to_consumer_share,
         `Close to \nFOOD HUB`= hub_proximity,
         `Close to \nFarmer's \nMarket` = d2c,
         `Food hub Intermediated Share` = foodhub_share,
         `SPMA` = spma,
         `Fruit Nut Share` = fruit_nut_share) %>%
  ggradar(font.radar = "roboto",
          values.radar   = c("0", "50", "100"),
          group.colours  = c("#013e75", "#a42423"),
          grid.label.size = 5,
          axis.label.size = 4,
          axis.label.offset = 1.2,
          gridline.min.colour = "#BFDBFE",
          gridline.mid.colour = "#60A5FA",
          gridline.max.colour = "#1E3A8A") +
  theme_void(base_family = "body_font")  +
  theme(
    plot.background = element_rect(
      fill = "#FAF8F2",
      color = NA
    ),
    panel.background = element_rect(
      fill = "#FAF8F2",
      color = NA
    ),

    plot.title = element_text(
      family = "title_font",
      size = 25,
      lineheight = 0.88,
      color = "#123B65",
      face = "plain",
      hjust = 0,
      margin = margin(b = 10)
    ),

    plot.subtitle = element_text(
      family = "body_font",
      size = 10,
      lineheight = 1.2,
      color = "#4B5D6D",
      hjust = 0,
      margin = margin(b = 12)
    ),

    legend.position = "right",
    legend.direction = "vertical",
    legend.key.width = unit(2, "cm"),
    legend.key.height = unit(0.14, "cm"),

    plot.margin = margin(
  t = 10,
  r = 10,
  b = 10,
  l = 10
)
) + coord_equal(clip = "off") 

fig_2_yakima

```

add the figure for arrows between the figures
```{r}
arrow_plot_01 <- ggplot() +
  annotate(
    "segment",
    x = 0.35, xend = 0.58,   # shorter arrow
    y = 0.62, yend = 0.62,
    linewidth = 1.3,
    color = "#123B65",
    arrow = arrow(
      type = "closed",
      length = unit(0.12, "inches")
    )
  ) +
  annotate(
    "text",
    x = 0.5,
    y = 0.35,
    label = "Moving food hubs closer to Lee\nimproves market access",
    family = "body_font",
    size = 4,
    lineheight = 1.1,
    color = "grey30",
    hjust = 0.5
  ) +
  xlim(0, 1) +
  ylim(0, 1) +
  theme_void() +
  theme(
    plot.background = element_rect(
      fill = "#faf8f2",
      color = NA
    ),
    panel.background = element_rect(
      fill = "#faf8f2",
      color = NA
    )
  )

arrow_plot_02 <- ggplot() +
  annotate(
    "segment",
    x = 0.35, xend = 0.58,   # shorter arrow
    y = 0.62, yend = 0.62,
    linewidth = 1.3,
    color = "#123B65",
    arrow = arrow(
      type = "closed",
      length = unit(0.12, "inches")
    )
  ) +
  annotate(
    "text",
    x = 0.5,
    y = 0.35,
    label = "Moving food hubs closer to Yakima\nimproves market access",
    family = "body_font",
    size = 4,
    lineheight = 1.1,
    color = "grey30",
    hjust = 0.5
  ) +
  xlim(0, 1) +
  ylim(0, 1) +
  theme_void() +
  theme(
    plot.background = element_rect(
      fill = "#faf8f2",
      color = NA
    ),
    panel.background = element_rect(
      fill = "#faf8f2",
      color = NA
    )
  )

```


wrap  the plot
```{r echo=FALSE}
fig_02 <- wrap_plots(fig_2_Greenville_Lee,arrow_plot_01, fig_2_lee, fig_2_Yakima_Okanogan,arrow_plot_02, fig_2_yakima, ncol = 3) +
  plot_annotation(
    title = "From Orchard to Opportunity",
    subtitle = "Food Hubs and Small Farm Market Access at the County Level",
    theme = theme(
      plot.background = element_rect(
        fill = "#FAF8F2",
        color = NA
      ),

      plot.title = element_text(
        family = "title_font",
        size = 25,
        lineheight = 0.88,
        color = "#123B65",
        face = "plain",
        hjust = 0.5,
        margin = margin(b = 10)
      ),

      plot.subtitle = element_text(
        family = "body_font",
        size = 10,
        lineheight = 1.2,
        color = "#4B5D6D",
        hjust = 0.5,
        margin = margin(b = 12)
      ),

      plot.margin = margin(
        t = 10,
        r = 10,
        b = 10,
        l = 10
      )
    )
  )
  

```
```{r}
ggsave("../figures/fig_02.pdf", plot = fig_02, width = 20, height = 8)

```